### 2025-02-13 Paper_Results.Rmd
First Commit

---
title: "Paper_Examples"
output: html_document
date: "2025-01-20"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(randomForest)
library(R.matlab)
library(foreach)
library(doParallel)
library(dplyr)
library(lme4)
library(tseries)
library(caret)
library(pROC)
library(boot)
library(patchwork)
```

## Pitfall 1: Within person performance

```{r }
############ Table 3 #############
##################################

# Basic Parameters from Study
n_features <- 10 # FIXED IN SAEB
n_samples <- 90  # Timepoints/ RECORDS (fixed in saeb) Number of samples per subject
n_subjects <- 150 # SUBJECTS (VARIED 4 TO 32) # Needs to be an even number

# Generating Outcome
overall_prob_outcome <- 0.1  # Overall probability of 1 (for the entire dataset) (e.g., tracking depression 0.146)
sqrt(overall_prob_outcome*(1-overall_prob_outcome))
sd_outcome <- 0.2 # Controls variability BETWEEN different subject (e.g., tracking depression 0.21)

# Generating Features
A <- 0.05 # Relationship between features and outcome
feature_std <- 0.1 # population level feature generating process
B <- 0.8  # Cross-subject variability ("random effect") (added per participants for all timepoints)
C <- 0.2 # Within-subject variability (added within participant for each timepoint)

source("/Users/f007qrc/Library/CloudStorage/GoogleDrive-anna.m.langener@dartmouth.edu/My Drive/Darmouth Drive/ML CrossValidation Project/Simulation_Functions_notshiny.R")

set.seed(12361488)
features_sample <- create_data(n_features,n_samples,n_subjects,A,feature_std,B,C,overall_prob_outcome,sd_outcome,time_effect)
features_sample <- features_sample[[1]]

sim_1 <- run_simulation(features_sample,"row-wise",1, testsize = 0.3)
sim_2 <- run_simulation_centering(features_sample,"row-wise",1,testsize = 0.3)
sim_3 <- run_simulation(features_sample,"subject-wise",1, testsize = 0.3)
sim_4 <- run_simulation_slidingwindow(features_sample,1,windowsize = 14)

table <- rbind(sim_1$overall_summary,sim_3$overall_summary, sim_4$overall_summary)
table$'AUC Overall' <- c(sim_1$auc_value,sim_3$auc_value,sim_4$auc_value)
colnames(table) <-  c("Mean AUC within-person:", "SD AUC within-person:", "% of AUC > 0.5 within-person:", "N included within-person:", "AUC Overall")
rownames(table) <- c("row-wise","subject-wise","rolling-window")
table <- round(table[,c(5,1,2,3,4)], 2)
table
```

```{r}
############# Figure 4 #################
########################################

result <- read.csv("/Users/f007qrc/Library/CloudStorage/GoogleDrive-anna.m.langener@dartmouth.edu/My Drive/Darmouth Drive/ML CrossValidation Project/simulation_results.csv")

p1 <- ggplot(result[result$A == 0,], aes(x=icc, y=auc, color = icc_pred)) +
  geom_point(alpha=0.4, size = 0.8) +
  theme_minimal() +
  scale_colour_gradientn(colours = c("#2A363B","#83AF9B","#C8C8A9","#F9CDAD","#FC9D9A","#FE4365")) +
  ylab("Prediction Performance (Overall AUC)") +
  xlab("% of Total Variation Explained by Between Person Differences in the Outcome \n(ICC Outcome)") +
  guides(col = guide_colourbar(title = "% of Total Variation Explained by \nBetween Person Differences \nin Predictors (ICC Predictor)")) +
  ggtitle("Bias in Performance Estimates (AUC) When No Relationship Exists: Expected AUC = 0.5") +
  geom_hline(yintercept = 0.5, color = "#83AF9B", size = 0.8) +
  annotate("text", x = 0.07, y = 0.82, label =  "Greater bias when between person differences explain a \nlarger % of total variation in both outcome and predictors.",  hjust = 0) +
  geom_curve(
    aes(x = 0.4, y = 0.85, xend = 0.68, yend = 0.87), 
    arrow = arrow(length = unit(0.08, "inch")), 
    size = 0.3, 
    color = "gray20", 
    curvature = -0.2
  ) + 
  annotate("text", x = 0.38, y = 0.54, label = "Low bias if between person differences explain a small \n% of total variation in outcome, predictors, or both.",  hjust = 0) +
    geom_curve(
    aes(x = 0.37, y = 0.54, xend = 0.33, yend = 0.51), 
    arrow = arrow(length = unit(0.08, "inch")), 
    size = 0.3, 
    color = "gray20", 
    curvature = 0.1
  ) +
  # Adding a diagonal line to illustrate the linear relationship
  geom_abline(slope = 0.5, intercept = 0.52, color = "#FE4365",  size = 0.8)

ggsave(  "/Users/f007qrc/Library/CloudStorage/GoogleDrive-anna.m.langener@dartmouth.edu/My Drive/Darmouth Drive/ML CrossValidation Project/Example2.jpg",
  plot = p1,
  width = 10.2,
  height = 4.5)

p1

mean(result$auc_individual[result$A == 0])
sd(result$auc_individual[result$A == 0])
mean(result$auc_c[result$A == 0])
sd(result$auc_c[result$A == 0])
sd(result$auc_individual[result$A == 0])

```

```{r}
############# Figure 5 #################
########################################
result <- read.csv("/Users/f007qrc/Library/CloudStorage/GoogleDrive-anna.m.langener@dartmouth.edu/My Drive/Darmouth Drive/ML CrossValidation Project/simulation_results.csv")

result <- result[result$A != 0 & result$A != 0.8,]

p3 <- ggplot(result, aes(x = icc, y = auc_c_individual - auc_individual)) +
  geom_point(aes(color = ifelse(auc_c_individual - auc_individual > 0, "Above Zero", "Below or Equal to Zero")), 
             alpha = 0.3, size = 0.8, show.legend = FALSE) +
  scale_color_manual(values = c("Above Zero" = "#FC9D9A", "Below or Equal to Zero" = "#83AF9B")) +
  theme_minimal() +
  ylab("Performance Difference \n(centered - uncentered)") +
  xlab("% of Total Variation Explained by Between Person Differences in the Outcome (ICC Outcome)") +
  ggtitle("Within-Person Performance") +
  geom_hline(yintercept = 0) + 
  annotate("text", x = 0, y = 0.28, label = "Centered predictors perform better in most cases (within-person)", hjust = 0) +
  annotate("text", x = 0, y = -0.08, label = "Uncentered predictors perform worse in most cases (within-person)", hjust = 0) +
  ylim(c(-0.45,0.35)) +
  geom_curve(
    aes(x = -0.01, y = 0.01, xend = -0.01, yend = 0.25), 
    arrow = arrow(length = unit(0.08, "inch")), 
    size = 0.5, 
    color = "#FC9D9A", 
    curvature = 0
  ) +
    geom_curve(
    aes(x = -0.01, y = 0.01, xend = -0.01, yend = -0.1), 
    arrow = arrow(length = unit(0.08, "inch")), 
    size = 0.5, 
    color = "#83AF9B", 
    curvature = 0
  )

p4 <- ggplot(result, aes(x = icc, y = auc_c - auc)) +
  geom_point(aes(color = ifelse(auc_c - auc > 0, "Above Zero", "Below or Equal to Zero")), 
             alpha = 0.3, size = 0.8,show.legend = FALSE) +
  scale_color_manual(values = c("Above Zero" = "#FC9D9A", "Below or Equal to Zero" = "#83AF9B")) +
  theme_minimal() +
  ylab("Performance Difference \n(centered - uncentered)") +
  xlab("% of Total Variation Explained by Between Person Differences in the Outcome (ICC Outcome)") +
  ggtitle("") +
  geom_hline(yintercept = 0) + 
  annotate("text", x = 0, y = 0.28, label = "Centered predictors perform better when between person differences explain a small % of total variance (overall)", hjust = 0) +
  annotate("text", x = 0.01, y = -0.39, label = "Uncentered predictors perform better when between person differences explain a large % of total variance (overall)", hjust = 0) +
  ylim(c(-0.45,0.35)) +
   ggtitle("Overall Performance (Within and Between-Person)")  +
    geom_curve(
    aes(x = -0.01, y = 0.01, xend = -0.01, yend = 0.27), 
    arrow = arrow(length = unit(0.08, "inch")), 
    size = 0.5, 
    color = "#FC9D9A", 
    curvature = 0
  ) +
    geom_curve(
    aes(x = 0.8, y = 0.01, xend = 0.8, yend = -0.41), 
    arrow = arrow(length = unit(0.08, "inch")), 
    size = 0.5, 
    color = "#83AF9B", 
    curvature = 0
  )

combined_plot <- (p3 / p4) +
  plot_annotation(
    title = "Performance difference: centered vs. not-centered",
  ) 
combined_plot

ggsave(  "/Users/f007qrc/Library/CloudStorage/GoogleDrive-anna.m.langener@dartmouth.edu/My Drive/Darmouth Drive/ML CrossValidation Project/Example4.jpg",
  plot = combined_plot,
  width = 10.2,
  height = 5.5)

mean(result$auc)
mean(result$auc_c)

mean(result$auc_c_individual)
mean(result$auc_individual)

test <- result$auc_c-result$auc
sum((test > 0)/length(test))

unique(result$A)
```

```{r}
############# Figure 6 #################
########################################
result <- read.csv("/Users/f007qrc/Library/CloudStorage/GoogleDrive-anna.m.langener@dartmouth.edu/My Drive/Darmouth Drive/ML CrossValidation Project/simulation_results.csv")
result <- result[result$A != 0 & result$A != 0.8,]
p1 <- ggplot(result, aes(x=auc_c, y=auc_c_individual, color = icc_pred)) +
  geom_point(alpha=0.5, size = 1) +
  #scale_size(range = c(.001, 6), name="") +
  theme_minimal() +
  scale_colour_gradientn(colours = c("#2A363B","#83AF9B","#C8C8A9","#F9CDAD","#FC9D9A","#FE4365")) +
  ylab("AUC within-person centered") +
  xlab("AUC centered") +
  guides(col = guide_colourbar(title = "% of Total Variation Explained by \nBetween Person Differences \nin Predictors (ICC Predictor)")) +
  geom_hline(yintercept = 0.5)   + 
annotate("text", x = 0.5, y = 0.9, label = "Linear relationship between overall AUC \nand within-person AUC",  hjust = 0) +
  geom_curve(
    aes(x = 0.63, y = 0.87, xend = 0.75, yend = 0.8), 
    arrow = arrow(length = unit(0.08, "inch")), 
    size = 0.3, 
    color = "gray20", 
    curvature = -0.15
  ) +
  ggtitle("Uncentered Predictors")

p2 <- ggplot(result, aes(x=auc, y=auc_individual, color = icc_pred)) +
  geom_point(alpha=0.5, size = 1) +
  #scale_size(range = c(.001, 6), name="") +
  theme_minimal() +
  scale_colour_gradientn(colours = c("#2A363B","#83AF9B","#C8C8A9","#F9CDAD","#FC9D9A","#FE4365")) +
  ylab("AUC within-person") +
  xlab("AUC") +
  guides(col = guide_colourbar(title = "% of Total Variation Explained by \nBetween Person Differences \nin Predictors (ICC Predictor)")) +
  geom_hline(yintercept = 0.5) + 
annotate("text", x = 0.5, y = 0.9, label = "If between person differences explain a large % of variation in predictors and outcome,\nthe overall AUC becomes high while the within-person AUC remains low.",  hjust = 0) +
  geom_curve(
    aes(x = 0.75, y = 0.82, xend = 0.9, yend = 0.6), 
    arrow = arrow(length = unit(0.08, "inch")), 
    size = 0.3, 
    color = "gray20", 
    curvature = 0.35
  ) +
  ggtitle("Centered Predictors")

combined_plot <- (p2 / p1) +
  plot_annotation(
    title = "Relationship Overall AUC and within-person AUC (A = 0.05 and A = 0.2)",
  ) 
combined_plot

ggsave(  "/Users/f007qrc/Library/CloudStorage/GoogleDrive-anna.m.langener@dartmouth.edu/My Drive/Darmouth Drive/ML CrossValidation Project/Example3.jpg",
  plot = combined_plot,
  width = 11.5,
  height = 6.5)
```

## Pitfall 2: Mismatch between chosen cross-validation strategy, employment, and actual observations 
```{r }
############ Table 7 #############
##################################

# Merge the p-values with the original dataset for plotting
data <- features_sample

# Compute mean y for each subject and reorder factor levels in ascending order
subject_stats <- data %>%
  group_by(subject) %>%
  summarize(mean_y = mean(y, na.rm = TRUE),
            has_y1 = any(y == 1)) %>%  # Identify if subject has any y = 1
  arrange(mean_y)

# Convert subject to a factor with the same levels as in data
subject_stats$subject <- factor(subject_stats$subject, levels = subject_stats$subject)

# Count the number of subjects with at least one y = 1
num_subjects_with_y1 <- sum(subject_stats$has_y1)

# Apply the new factor levels to data
data <- data %>%
  mutate(subject = factor(subject, levels = subject_stats$subject))

# Merge "has_y1" information into the main dataset
data <- left_join(data, subject_stats, by = "subject")

# Create the sorted plot with conditional background color
p2 <- ggplot(data, aes(x = time, y = y)) +
  # Add background shading for subjects with y = 1
  geom_rect(
    data = data %>% filter(!has_y1),
    aes(xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf),
    fill = "#E5E4E2", alpha = 0.05, inherit.aes = T
  ) +
  geom_point(color = "#FC9D9A", size = 0.5) +                       
  facet_wrap(~subject) +               
  labs(
    x = "Time", 
    y = "Outcome", 
    title = paste("Within-person variability over time: \n88 participants (58.6%) do not experience within-person variability. ")
  ) +
  scale_y_continuous(breaks = c(0, 1)) +  # Ensure only 0 and 1 appear on the y-axis
  theme_minimal() +
  theme(axis.text.x=element_blank(),axis.ticks.x=element_blank()) +
  guides(fill = "none")

p2

ggsave(  "/Users/f007qrc/Library/CloudStorage/GoogleDrive-anna.m.langener@dartmouth.edu/My Drive/Darmouth Drive/ML CrossValidation Project/Example_var.jpg",
  plot = p2,
  width = 8,
  height = 7)

```

## Pitfall 3: Weak benchmark and missing baseline comparison

```{r }
############ Table 5 #############
##################################

sim_1$auc_value_base # row-wise
sim_4$auc_value_base # rolling window

sim_4$auc_value

test <- sim_1$ind_base[[1]] %>%
  group_by(subject) %>%
  filter(length(unique(true)) > 1) %>%
  filter(sum(true == 1) > 0, sum(true == 0) > 0) %>%
  summarise(
    auc_val = auc(roc(as.numeric(as.character(true)), as.numeric(as.character(pred)), quiet = TRUE))[1],
    .groups = "drop" # Ungroup after summarizing
  )

mean(test$auc_val)

```

```{r }
############ Figure 8 #############
##################################

result = read.csv("/Users/f007qrc/Library/CloudStorage/GoogleDrive-anna.m.langener@dartmouth.edu/My Drive/Darmouth Drive/ML CrossValidation Project/simulation_results_baseline.csv")
p7 <- ggplot(result, aes(x=icc, y=auc_value_base)) +
  geom_point(alpha=0.5, size = 1.2, color = "#83AF9B") +
  #scale_size(range = c(.001, 6), name="") +
  theme_minimal() +
  scale_colour_gradientn(colours = c("#83AF9B")) +
  ylab("Overall Performance (AUC)") +
  xlab("% of Total Variation Explained by Between Person Differences in the Outcome (ICC Outcome)") +
  guides(col = guide_colourbar()) +
  ggtitle(paste("Performance of a Random Intercept Only Model (No Predictors Included)")) +
  geom_hline(yintercept = 0.5) +
  ylim(range= c(0.5,1)) +
    annotate("text", x = 0.5, y = 0.82, label =  "Performance increases as between person differences explain a \nlarger % of total variation in the outcome",  hjust = 0) +
  geom_curve(
    aes(x = 0.7, y = 0.85, xend = 0.5, yend = 0.91), 
    arrow = arrow(length = unit(0.08, "inch")), 
    size = 0.3, 
    color = "gray20", 
    curvature = 0.2
  ) +
      annotate("text", x = 0.2, y = 0.62, label =  "Performance can be moderate even if between person differences explain only a relatively \nsmall % of total variation in the outcome",  hjust = 0) +
  geom_curve(
    aes(x = 0.19, y = 0.62, xend = 0.07, yend = 0.61), 
    arrow = arrow(length = unit(0.08, "inch")), 
    size = 0.3, 
    color = "gray20", 
    curvature = 0.2
  )

ggsave(  "/Users/f007qrc/Library/CloudStorage/GoogleDrive-anna.m.langener@dartmouth.edu/My Drive/Darmouth Drive/ML CrossValidation Project/figure7.jpg",
  plot = p7,
  width = 10,
  height = 5)

```

### 2025-02-13 ExtensionSaebv2.R

@@ -1,0 +1,879 @@
+######### Load Packages ############
+####################################
+
+library(randomForest)
+library(R.matlab)
+library(foreach)
+library(doParallel)
+library(dplyr)
+library(lme4)
+library(tseries)
+library(caret)
+library(pROC)
+library(boot)
+library(plotly)
+
+
+#set.seed(1234)  # Setting the seed for random number generation
+##################################################
+########### Set Simualtion Parameters ############
+##################################################
+# Basic Parameters from Study
+n_features <- 10 # FIXED IN SAEB
+n_samples <- 90  # Timepoints/ RECORDS (fixed in saeb) Number of samples per subject
+n_subjects <- 150 # SUBJECTS (VARIED 4 TO 32) # Needs to be an even number
+
+# Generating Outcome
+overall_prob_outcome <- 0.3  # Overall probability of 1 (for the entire dataset) (e.g., tracking depression 0.146)
+sqrt(overall_prob_outcome*(1-overall_prob_outcome))
+sd_outcome <- 0.05 # Controls variability BETWEEN different subject (e.g., tracking depression 0.21)
+# within_variability_outcome has to be: sd_outcome < sqrt(overall_prob_outcome*(1-overall_prob_outcome))
+# (variance < mean(1-mean)
+
+time_effect = FALSE
+
+# Generating Features
+A <- 0.05 # Relationship between features and outcome
+feature_std <- 0.1 # population level feature generating process
+B <- 0.8  # Cross-subject variability ("random effect") (added per participants for all timepoints)
+# Add varying B (B[i], C[i] should work)
+C <- 0.1 # Within-subject variability (added within participant for each timepoint)
+
+##################################################
+################# Data Simulation ################
+##################################################
+source("/Users/f007qrc/Library/CloudStorage/GoogleDrive-anna.m.langener@dartmouth.edu/My Drive/Darmouth Drive/ML CrossValidation Project/Simulation_Functions_notshiny.R")
+
[…]

### 2025-02-13 Simulation_Functions_notshiny.R

@@ -1,0 +1,588 @@
+##############################
+### Simulation Functions #####
+##############################
+
+estBetaParams <- function(mu, var) {
+  alpha <- ((1 - mu) / var - 1 / mu) * mu ^ 2
+  beta <- alpha * (1 / mu - 1)
+  return(params = list(alpha = alpha, beta = beta))
+} #taken from https://stats.stackexchange.com/questions/12232/calculating-the-parameters-of-a-beta-distribution-using-the-mean-and-variance/12239#12239
+
+create_data <- function(n_features,n_samples,n_subjects,A,feature_std,B,C,overall_prob_outcome,sd_outcome,time_effect){
+  
+  y <- vector("numeric", length = n_subjects * n_samples) 
+  subject_id <- rep(1:n_subjects, each = n_samples)  # Repeat subject IDs for each timepoint
+  time_id <- rep(1:n_samples, times = n_subjects)  # Repeat timepoints for each subject
+  data <- data.frame(subject = subject_id, time = time_id, y = y)
+  
+  # Simulate Outcome
+  para <- estBetaParams(overall_prob_outcome,sd_outcome^2)
+  prob_l <- rbeta(n=n_subjects ,shape1 = para$alpha, shape2 = para$beta) # sample probability for each subject
+  
+  print(paste("Mean prob:",mean(prob_l)))
+  print(paste("SD prob:",sd(prob_l)))
+  
+  prob_l = rep(prob_l, each = n_samples) 
+  
+  if (time_effect) {
+    # NEEDS TO BE ADDED
+   # t_effect = seq(-1, 1, length.out = n_samples)
+    #t_effect = rep(t_effect, times = n_subjects)
+    #prob_matrix = 1/(1+exp(-(prob_l+t_effect)))
+  } else {
+    prob_matrix = prob_l
+  }
+  
+  data$y <- rbinom(n_subjects * n_samples, size = 1, prob = prob_matrix) #sample 0 or 1 for each subject & timepoint, prob_matrix indicates the probability for each subject
+  data$A = data$y
+
+  # Simulate Features (adapted from Saeb et al. and https://aosmith.rbind.io/2018/04/23/simulate-simulate-part-2/)
+  data$A[data$y == 0] = -A # relationship to outcome
+  data$A[data$y == 1] = A
+  
+  features_sample <- list()
+  
[…]

### 2025-02-13 Paper_Results.Rmd
update files

@@
 ---
-title: "Paper_Examples"
+title: "Paper_Results"
 output: html_document
@@
 library(randomForest)
-library(R.matlab)
-library(foreach)
-library(doParallel)
 library(dplyr)
 library(lme4)
-library(tseries)
 library(caret)
 library(pROC)
-library(boot)
 library(patchwork)
 ```
+## General Information
+This R Markdown file can be used to replicate our results that are presented in the paper. Please note that it does not contain the shiny app.
 
@@
 # Basic Parameters from Study
-n_features <- 10 # FIXED IN SAEB
-n_samples <- 90  # Timepoints/ RECORDS (fixed in saeb) Number of samples per subject
-n_subjects <- 150 # SUBJECTS (VARIED 4 TO 32) # Needs to be an even number
+n_features <- 10  # Features
+n_samples <- 90  # Timepoints
+n_subjects <- 150 # Subjects
 
@@
 overall_prob_outcome <- 0.1  # Overall probability of 1 (for the entire dataset) (e.g., tracking depression 0.146)
-sqrt(overall_prob_outcome*(1-overall_prob_outcome))
-sd_outcome <- 0.2 # Controls variability BETWEEN different subject (e.g., tracking depression 0.21)
+sd_outcome <- 0.2 # Controls variability BETWEEN different subject (nneds to be smaller than sqrt(overall_prob_outcome*(1-overall_prob_outcome)))
 
@@
 
-source("/Users/f007qrc/Library/CloudStorage/GoogleDrive-anna.m.langener@dartmouth.edu/My Drive/Darmouth Drive/ML CrossValidation Project/Simulation_Functions_notshiny.R")
+source("Simulation_Functions_notshiny.R")
 
 set.seed(12361488)
-features_sample <- create_data(n_features,n_samples,n_subjects,A,feature_std,B,C,overall_prob_outcome,sd_outcome,time_effect)
+features_sample <- create_data(n_features,n_samples,n_subjects,A,feature_std,B,C,overall_prob_outcome,sd_outcome,time_effect = FALSE)
 features_sample <- features_sample[[1]]
@@
 
-result <- read.csv("/Users/f007qrc/Library/CloudStorage/GoogleDrive-anna.m.langener@dartmouth.edu/My Drive/Darmouth Drive/ML CrossValidation Project/simulation_results.csv")
+result <- read.csv("simulation_results.csv") # see Run_Simulation.R to investigate how results were created
 
@@
 
-ggsave(  "/Users/f007qrc/Library/CloudStorage/GoogleDrive-anna.m.langener@dartmouth.edu/My Drive/Darmouth Drive/ML CrossValidation Project/Example2.jpg",
+ggsave("Figures/Example2.jpg",
   plot = p1,
@@
 ########################################
-result <- read.csv("/Users/f007qrc/Library/CloudStorage/GoogleDrive-anna.m.langener@dartmouth.edu/My Drive/Darmouth Drive/ML CrossValidation Project/simulation_results.csv")
+result <- read.csv("simulation_results.csv")
 
@@
 
-ggsave(  "/Users/f007qrc/Library/CloudStorage/GoogleDrive-anna.m.langener@dartmouth.edu/My Drive/Darmouth Drive/ML CrossValidation Project/Example4.jpg",
+ggsave(  "Figures/Example4.jpg",
   plot = combined_plot,
@@
 
-mean(result$auc)
-mean(result$auc_c)
-
-mean(result$auc_c_individual)
-mean(result$auc_individual)
-
-test <- result$auc_c-result$auc
-sum((test > 0)/length(test))
-
-unique(result$A)
 ```
@@
 ########################################
-result <- read.csv("/Users/f007qrc/Library/CloudStorage/GoogleDrive-anna.m.langener@dartmouth.edu/My Drive/Darmouth Drive/ML CrossValidation Project/simulation_results.csv")
+result <- read.csv("simulation_results.csv")
 result <- result[result$A != 0 & result$A != 0.8,]
@@
 
-ggsave(  "/Users/f007qrc/Library/CloudStorage/GoogleDrive-anna.m.langener@dartmouth.edu/My Drive/Darmouth Drive/ML CrossValidation Project/Example3.jpg",
+ggsave( "Figures/Example3.jpg",
   plot = combined_plot,
@@
 
-ggsave(  "/Users/f007qrc/Library/CloudStorage/GoogleDrive-anna.m.langener@dartmouth.edu/My Drive/Darmouth Drive/ML CrossValidation Project/Example_var.jpg",
+ggsave( "Figures/Example_var.jpg",
   plot = p2,
@@
 ##################################
-
 sim_1$auc_value_base # row-wise
@@
 
-result = read.csv("/Users/f007qrc/Library/CloudStorage/GoogleDrive-anna.m.langener@dartmouth.edu/My Drive/Darmouth Drive/ML CrossValidation Project/simulation_results_baseline.csv")
+result = read.csv("simulation_results_baseline.csv")
 p7 <- ggplot(result, aes(x=icc, y=auc_value_base)) +
@@
 
-ggsave(  "/Users/f007qrc/Library/CloudStorage/GoogleDrive-anna.m.langener@dartmouth.edu/My Drive/Darmouth Drive/ML CrossValidation Project/figure7.jpg",
+ggsave(  "Figures/figure7.jpg",
   plot = p7,

### 2025-02-13 ExtensionSaebv2.R

@@ -1,879 +1,0 @@
-######### Load Packages ############
-####################################
-
-library(randomForest)
-library(R.matlab)
-library(foreach)
-library(doParallel)
-library(dplyr)
-library(lme4)
-library(tseries)
-library(caret)
-library(pROC)
-library(boot)
-library(plotly)
-
-
-#set.seed(1234)  # Setting the seed for random number generation
-##################################################
-########### Set Simualtion Parameters ############
-##################################################
-# Basic Parameters from Study
-n_features <- 10 # FIXED IN SAEB
-n_samples <- 90  # Timepoints/ RECORDS (fixed in saeb) Number of samples per subject
-n_subjects <- 150 # SUBJECTS (VARIED 4 TO 32) # Needs to be an even number
-
-# Generating Outcome
-overall_prob_outcome <- 0.3  # Overall probability of 1 (for the entire dataset) (e.g., tracking depression 0.146)
-sqrt(overall_prob_outcome*(1-overall_prob_outcome))
-sd_outcome <- 0.05 # Controls variability BETWEEN different subject (e.g., tracking depression 0.21)
-# within_variability_outcome has to be: sd_outcome < sqrt(overall_prob_outcome*(1-overall_prob_outcome))
-# (variance < mean(1-mean)
-
-time_effect = FALSE
-
-# Generating Features
-A <- 0.05 # Relationship between features and outcome
-feature_std <- 0.1 # population level feature generating process
-B <- 0.8  # Cross-subject variability ("random effect") (added per participants for all timepoints)
-# Add varying B (B[i], C[i] should work)
-C <- 0.1 # Within-subject variability (added within participant for each timepoint)
-
-##################################################
-################# Data Simulation ################
-##################################################
-source("/Users/f007qrc/Library/CloudStorage/GoogleDrive-anna.m.langener@dartmouth.edu/My Drive/Darmouth Drive/ML CrossValidation Project/Simulation_Functions_notshiny.R")
-
[…]

### 2025-02-13 Run_Simulation.R

@@ -1,0 +1,220 @@
+####################################
+######### Load Packages ############
+####################################
+
+library(randomForest)
+library(foreach)
+library(doParallel)
+library(dplyr)
+library(lme4)
+library(caret)
+library(pROC)
+
+####################################################
+################ Run Simulation ####################
+####################################################
+source("Simulation_Functions_notshiny.R")
+
+n_features <- 10 
+n_samples <- 90  
+n_subjects <- 150 
+time_effect = FALSE
+
+A <- 0 # Relationship between features and outcome, we also ran this for A = 0.05 and A = 0.2, and then merged the results
+feature_std <- 0.1 # population level feature generating process
+
+# Initialize parameters
+overall_prob_outcome <- seq(0.1, 0.9, by = 0.1)  
+overall_prob_outcome <- 0.5
+sd_outcome <- seq(0.05, 0.29, by = 0.05)            
+B <- seq(0.1, 0.9, by = 0.05)       
+C <-  seq(0.05,0.45, by = 0.05)   
+
+
+# Create a grid of all parameter combinations
+param_grid <- expand.grid(overall_prob_outcome = overall_prob_outcome,
+                          sd_outcome = sd_outcome,
+                          B = B,
+                          C = C)
+
+# Set up parallel backend
+n_cores <- parallel::detectCores() - 1  
+cl <- makeCluster(n_cores)
+registerDoParallel(cl)
+old <- Sys.time() # get start time
+
+# Parallel loop with foreach
+result <- foreach(i = 1:nrow(param_grid), .combine = rbind, .packages = c("dplyr", "lme4", "pROC", "randomForest"), .errorhandling = "remove") %dopar% {
+  # Extract parameter combination for this iteration
+  params <- param_grid[i, ]
+  overall_prob <- params$overall_prob_outcome
+  sd <- params$sd_outcome
+  B_value <- params$B
+  C_value <- params$C
+  
+  # Generate data for this iteration
+  features_sample <- create_data(n_features, n_samples, n_subjects, A, feature_std, B_value, C_value, overall_prob, sd, time_effect)
+  features_sample_Astd <- features_sample[[2]]
+  features_sample <- features_sample[[1]]
+  
+  ### Descriptives ####
+  test <- features_sample %>% group_by(subject) %>% summarise(prob = mean(y))
[…]

### 2025-02-13 Simulation_Functions_notshiny.R

@@ -25,10 +25,7 @@ create_data <- function(n_features,n_samples,n_subjects,A,feature_std,B,C,overal
   prob_l = rep(prob_l, each = n_samples) 
   
   if (time_effect) {
-    # NEEDS TO BE ADDED
-   # t_effect = seq(-1, 1, length.out = n_samples)
-    #t_effect = rep(t_effect, times = n_subjects)
-    #prob_matrix = 1/(1+exp(-(prob_l+t_effect)))
+    #Can be added in the future
   } else {
     prob_matrix = prob_l
   }
@@ -77,17 +74,6 @@ create_data <- function(n_features,n_samples,n_subjects,A,feature_std,B,C,overal
     }
     
   }
-  # Center the training set
-  # subject_means <- features_sample[[1]] %>%
-  #   group_by(subject) %>%
-  #   summarise(across(all_of(colnames(features_sample[[1]][,1:10])), mean), .groups = "drop")
-  # 
-  # # Center the training set
-  # features_sample[[3]] <- features_sample[[1]] %>%
-  #   left_join(subject_means, by = "subject", suffix = c("", "_mean")) %>%
-  #   mutate(across(all_of(colnames(features_sample[[1]][,1:10])), ~ . - get(paste0(cur_column(), "_mean")))) %>%
-  #   select(all_of(colnames(features_sample[[1]][,1:10])),"subject","time","y","A")
-  
   return(features_sample)
 }
 
@@ -283,44 +269,24 @@ run_simulation_centering <- function(features_sample,cv,n_bootstrap,testsize, se
     feature_names = colnames(features_sample)[1:10]
    
   
-    # # Center the training set
-    # subject_means <- features_sample[samples_train, ] %>%
-    #  group_by(subject) %>%
-    #  summarise(across(all_of(names(train_X)), mean), .groups = "drop")
-    # # 
-    # # subject_means <- data.frame(subject = unique(features_sample[[id_variable]]))
-    # # 
-    # # for(j in 1:n_features){
-    # #   subject_means$variable[j] <- feature_names[j]
-    # # 
-    # #   model_pred <- lmer(as.formula(paste0(feature_names[j], "~ 1 + ( 1  |", id_variable,")")), data = features_sample[samples_train,])
-    # #   # Extract the random effects for the current feature
-    # #   ranef_means <- as.data.frame(ranef(model_pred)$subject)
-    # # 
-    # #   # Add the feature-specific random effects to the subject_means data frame
[…]

### 2025-02-13 Simulation_Functions_notshiny.R
Update Simulation_Functions_notshiny.R

@@ -57,7 +57,7 @@ create_data <- function(n_features,n_samples,n_subjects,A,feature_std,B,C,overal
     timeeff = C * rnorm(n_subjects*n_samples, mean = 0, sd = 1) #C * rnorm(n_subjects*n_samples)
     maineff = features_sample[[1]]$A
     
-    features_sample[[1]][,i] = maineff + feature_noise + subjecteff + timeeff 
+    features_sample[[1]][,i] = maineff + feature_noise + subjecteff + timeeff  # ONlY THIS IS RELEVANT
     features_sample[[1]][, paste("V", i, "_mean", sep = "")] <- subjecteff
     features_sample[[2]][,i] = maineff + feature_noise + timeeff 
     
@@ -89,7 +89,7 @@ run_simulation <- function(features_sample,cv,n_bootstrap,testsize, seed = "1236
   acc <- numeric(n_bootstrap)
   auc_value <- numeric(n_bootstrap)
   acc_rep <- numeric(n_bootstrap)
-  true_list <- list()  # To store true_sw for each iteration
+  true_list <- list()  
   pred_list <- list() 
   subject <- list() 
   ind <- list()
@@ -97,7 +97,7 @@ run_simulation <- function(features_sample,cv,n_bootstrap,testsize, seed = "1236
   acc_base <- numeric(n_bootstrap)
   auc_value_base <- numeric(n_bootstrap)
   acc_rep_base <- numeric(n_bootstrap)
-  true_list_base <- list()  # To store true_sw for each iteration
+  true_list_base <- list()  
   pred_list_base <- list() 
   ind_base <- list()
   
@@ -166,7 +166,7 @@ run_simulation <- function(features_sample,cv,n_bootstrap,testsize, seed = "1236
       ind_base[[i]] = data.frame(subject = subject[[i]], true = true_list_base[[i]], pred = pred_list_base[[i]])
       }
     }
-  } # end bootstrap
+  } # end bootstrap (we don't use the bootstrap in the paper)
 
   if(cv == "row-wise"){ 
     print(paste("Baseline Mean AUC:",mean(auc_value_base)))
@@ -223,11 +223,12 @@ run_simulation <- function(features_sample,cv,n_bootstrap,testsize, seed = "1236
 } 
 ############################################
 run_simulation_centering <- function(features_sample,cv,n_bootstrap,testsize, seed = "12361488"){
+  # Same function as before but this time we center the variables. In the shiny app we also include a function to center variables for the rolling window.
[…]

### 2025-02-15 Paper_Results.Rmd
update figures

@@
 
-ggsave("Figures/Example2.jpg",
+ggsave("Figures/Figure4_Example2.jpg",
   plot = p1,
@@
 
-ggsave(  "Figures/Example4.jpg",
+ggsave(  "Figures/Figure5_Example4.jpg",
   plot = combined_plot,
@@
 
-ggsave( "Figures/Example3.jpg",
+ggsave( "Figures/Figure6_Example3.jpg",
   plot = combined_plot,
@@
 
-ggsave( "Figures/Example_var.jpg",
+ggsave( "Figures/Figure7_Example_var.jpg",
   plot = p2,
@@
 
-ggsave(  "Figures/figure7.jpg",
+ggsave(  "Figures/Figure8.jpg",
   plot = p7,

### 2025-06-10 Paper_Results.Rmd
Review Changes

@@
 
-## Pitfall 1: Within person performance
+## Pitfall 1: Within person performance Emprircal Example
+```{r }
+### Descriptives
+
+# Read the CSV files
+final_results_df_2_b <- read.csv("final_results_df_2_b.csv")
+final_results_df_3_b <- read.csv("final_results_df_3_b.csv")
+final_results_df_4_b <- read.csv("final_results_df_4_b.csv")
+final_results_df_5_b <- read.csv("final_results_df_5_b.csv")
+
+# Select specific columns
+cols_to_select1 <- c("outcome", "overall_prob_outcome", "sd_outcome",
+                  "icc_data")
+cols_to_select <- c("overall_prob_outcome", "sd_outcome",
+                   "icc_data")
+
+df2_selected <- final_results_df_2_b[, cols_to_select1]
+df3_selected <- final_results_df_3_b[, cols_to_select]
+df4_selected <- final_results_df_4_b[, cols_to_select]
+df5_selected <- final_results_df_5_b[, cols_to_select]
+
+# Rename columns to distinguish between dataframes (optional but recommended)
+names(df2_selected) <- paste0(names(df2_selected), "_df2")
+names(df3_selected) <- paste0(names(df3_selected), "_df3")
+names(df4_selected) <- paste0(names(df4_selected), "_df4")
+names(df5_selected) <- paste0(names(df5_selected), "_df5")
+
+# Combine the selected dataframes column-wise
+combined_df <- cbind(df2_selected, df3_selected, df4_selected, df5_selected)
+
+# View or write the combined data frame
+head(combined_df)
+
+combined_df <- combined_df %>%
+  mutate(across(where(is.numeric), ~ round(.x, 2)))
+write.csv(combined_df, "combined_des_binary.csv", row.names = FALSE)
+
+# Read the CSV files
+final_results_df_2 <- read.csv("final_results_df_2.csv")
+final_results_df_3 <- read.csv("final_results_df_3.csv")
+final_results_df_4 <- read.csv("final_results_df_4.csv")
+final_results_df_5 <- read.csv("final_results_df_5.csv")
+
+# Select specific columns
+cols_to_select1 <- c("outcome", "overall_prob_outcome", "sd_outcome",
+                  "icc_data")
+cols_to_select <- c("overall_prob_outcome", "sd_outcome",
+                   "icc_data")
+
+df2_selected <- final_results_df_2[, cols_to_select1]
+df3_selected <- final_results_df_3[, cols_to_select]
+df4_selected <- final_results_df_4[, cols_to_select]
+df5_selected <- final_results_df_5[, cols_to_select]
+
+# Rename columns to distinguish between dataframes (optional but recommended)
+names(df2_selected) <- paste0(names(df2_selected), "_df2")
+names(df3_selected) <- paste0(names(df3_selected), "_df3")
+names(df4_selected) <- paste0(names(df4_selected), "_df4")
+names(df5_selected) <- paste0(names(df5_selected), "_df5")
+
+# Combine the selected dataframes column-wise
+combined_df <- cbind(df2_selected, df3_selected, df4_selected, df5_selected)
+
+# View or write the combined data frame
+head(combined_df)
+
+combined_df <- combined_df %>%
+  mutate(across(where(is.numeric), ~ round(.x, 2)))
+write.csv(combined_df, "combined_des_continious.csv", row.names = FALSE)
+
+```
 
 ```{r }
-############ Table 3 #############
-##################################
+############ Table 4 #############
+####### Simulation Study #########
 
@@
 n_features <- 10  # Features
-n_samples <- 90  # Timepoints
+n_samples <- 200  # Timepoints
 n_subjects <- 150 # Subjects
@@
 overall_prob_outcome <- 0.1  # Overall probability of 1 (for the entire dataset) (e.g., tracking depression 0.146)
-sd_outcome <- 0.2 # Controls variability BETWEEN different subject (nneds to be smaller than sqrt(overall_prob_outcome*(1-overall_prob_outcome)))
+sd_outcome <- 0.25 # Controls variability BETWEEN different subject (nneds to be smaller than sqrt(overall_prob_outcome*(1-overall_prob_outcome)))
 
@@
 B <- 0.8  # Cross-subject variability ("random effect") (added per participants for all timepoints)
-C <- 0.2 # Within-subject variability (added within participant for each timepoint)
+C <- 0.45 # Within-subject variability (added within participant for each timepoint)
 
@@
 set.seed(12361488)
-features_sample <- create_data(n_features,n_samples,n_subjects,A,feature_std,B,C,overall_prob_outcome,sd_outcome,time_effect = FALSE)
+features_sample <- create_data(n_features,n_samples,n_subjects,A,feature_std,B,C,overall_prob_outcome,sd_outcome,0,0)
 features_sample <- features_sample[[1]]
@@
 
+```{r }
+############ Table 5 #############
+############# Binary #############
+
+source("Simulation_UploadData.R")
+data = read.csv("esm_clean.csv")
+set.seed(1234)
+
+data = data  %>% group_by %>%
+  group_by(user_id) %>%
+  mutate(day = row_number())
+
+data <- data %>%
+  group_by(user_id) %>%
+  filter(n_distinct(day) >= 150) %>%
+  ungroup()
+data = data[data$day <= 150,]
+
+sampled_ids <- sample(unique(data$user_id), size = 20, replace = FALSE)
+
+data <- data %>% 
+  filter(user_id %in% sampled_ids)
+
+data$user_id <- sapply(data$user_id, match, unique(unlist(data$user_id)))
+
+id_variable = "user_id"   
+time_variable = "day"
+
+colnames(data)[colnames(data) == "user_id"] <- "subject"
+colnames(data)[colnames(data) == "day"]     <- "time"
+#colnames(data)[colnames(data) == "wantedToFeel_goodAboutMe"]  <- "y"
+
+data <- data %>%
+  group_by(subject) %>%
+  mutate(y = lag(wantedToFeel_goodAboutMe, n = 1))
+
+data <- data[!is.na(data$y),]
+
+## Run strategies
+r_1 <- run_simulation_own(
+  features_sample = data,
+  cv              = "record-wise",
+  n_bootstrap     = 1,
+  testsize        = 0.3,
+  n_features      =  c("feelingBadToGood", "energy", 
+                       "thinkingOverAndOver", "lonely", "selfWorth", "appreciating", "stressed"),
+  seed = 1234
+
+)
+
+r_3 <- run_simulation_own(
+  features_sample = data,
+  cv              = "subject-wise",
+  n_bootstrap     = 1,
+  testsize        = 0.3,
+  n_features      =  c("feelingBadToGood", "energy", 
+                       "thinkingOverAndOver", "lonely", "selfWorth", "appreciating", "stressed"),
+  seed = 1234
+
+)
+r_4 <- run_simulation_slidingwindow_own(
+  features_sample = data,
+  n_bootstrap     = 1,
+  windowsize      = 14,
+  n_features      = c("feelingBadToGood", "energy",
+                      "thinkingOverAndOver", "lonely", "selfWorth", "appreciating", "stressed"),
+  seed = 1234
+)
+
+table <- rbind(r_1$overall_summary,r_3$overall_summary, r_4$overall_summary)
+table$'AUC Overall' <- c(r_1$auc_value,r_3$auc_value,r_4$auc_value)
+colnames(table) <-  c("Mean AUC within-person:", "SD AUC within-person:", "% of AUC > 0.5 within-person:", "N included within-person:", "AUC Overall")
+rownames(table) <- c("row-wise","subject-wise","rolling-window")
+table
+```
+
+```{r }
+############ Table 5 #############
+############# Continious #########
+
+source("Simulation_continous.R")
+
+data = read.csv("esm_clean.csv")
+set.seed(1234)
+
+data = data  %>% group_by %>%
+  group_by(user_id) %>%
+  mutate(day = row_number())
+
+data <- data %>%
+  group_by(user_id) %>%
+  filter(n_distinct(day) >= 150) %>%
+  ungroup()
+data = data[data$day <= 150,]
+
+sampled_ids <- sample(unique(data$user_id), size = 20, replace = FALSE)
+
+data <- data %>% 
+  filter(user_id %in% sampled_ids)
+
+data$user_id <- sapply(data$user_id, match, unique(unlist(data$user_id)))
+
+id_variable = "user_id"   
+time_variable = "day"
+
+colnames(data)[colnames(data) == "user_id"] <- "subject"
+colnames(data)[colnames(data) == "day"]     <- "time"
+#colnames(data)[colnames(data) == "wantedToFeel_goodAboutMe"]  <- "y"
+
+data <- data %>%
+  group_by(subject) %>%
+  mutate(y = lag(lonely, n = 1))
+
+data <- data[!is.na(data$y),]
+
+## Run strategies
+r_1 <- run_simulation_own(
+  features_sample = data,
+  cv              = "record-wise",
+  n_bootstrap     = 1,
+  testsize        = 0.3,
+  n_features      =  c("feelingBadToGood", "energy", 
+                       "thinkingOverAndOver", "selfWorth", "appreciating", "stressed"),
+  seed = 1234
+
+)
+
+r_3 <- run_simulation_own(
+  features_sample = data,
+  cv              = "subject-wise",
+  n_bootstrap     = 1,
+  testsize        = 0.3,
+  n_features      =  c("feelingBadToGood", "energy", 
+                       "thinkingOverAndOver", "selfWorth", "appreciating", "stressed"),
+  seed = 1234
+
+)
+r_4 <- run_simulation_slidingwindow_own(
+  features_sample = data,
+  n_bootstrap     = 1,
+  windowsize      = 14,
+  n_features      = c("feelingBadToGood", "energy",
+                      "thinkingOverAndOver", "selfWorth", "appreciating", "stressed"),
+  seed = 1234
+)
+
+table <- rbind(
+  r_1[c("rsq", "cor", "overall_summary.mean_r2", "overall_summary.mean_cor", "overall_summary.sd")],
+  r_3[c("rsq", "cor", "overall_summary.mean_r2", "overall_summary.mean_cor", "overall_summary.sd")],
+  r_4[c("rsq", "cor", "overall_summary.mean_r2", "overall_summary.mean_cor", "overall_summary.sd")]
+)
+
+table <- round(table[,c(5,1,2,3,4)], 2)
+table
+```
+
+```{r }
+## Try out
+
+# Basic Parameters from Study
+n_features <- 10  # Features
+n_samples <- 200  # Timepoints
+n_subjects <- 15 # Subjects
+
+# Generating Outcome
+overall_prob_outcome <- 0.4  # Overall probability of 1 (for the entire dataset) (e.g., tracking depression 0.146)
+sd_outcome <- 0.3 # Controls variability BETWEEN different subject (nneds to be smaller than sqrt(overall_prob_outcome*(1-overall_prob_outcome)))
+
+# Generating Features
+A <- 0.02 # Relationship between features and outcome
+feature_std <- 0.1 # population level feature generating process
+B <- 0.8  # Cross-subject variability ("random effect") (added per participants for all timepoints)
+C <- 0.2 # Within-subject variability (added within participant for each timepoint)
+
+source("Simulation_Functions_notshiny.R")
+
+set.seed(12361488)
+features_sample <- create_data(n_features, n_samples, n_subjects, A, feature_std, B, C, overall_prob_outcome, sd_outcome, 0,0)
+
+features_sample <- features_sample[[1]]
+
+sim_1 <- run_simulation(features_sample,"row-wise",1, testsize = 0.3)
+
+table <- rbind(sim_1$overall_summary)
+table$'AUC Overall' <- c(sim_1$auc_value)
+
+table
+
+library(dplyr)
+library(ggplot2)
+
+pred_true = as.data.frame(sim_1$ind)
+
+pred_true <- pred_true %>%
+  arrange(subject, true) %>%
+  group_by(subject) %>%
+  mutate(day = row_number())
+
+library(pROC)
+library(dplyr)
+
+# Compute individual AUC
+ind_auc <- pred_true %>%
+  group_by(subject) %>%
+  filter(length(unique(true)) > 1) %>%
+  filter(sum(true == 1) > 0, sum(true == 0) > 0) %>%
+  summarise(
+    auc_val = pROC::auc(pROC::roc(response = true, predictor = pred, direction = "<",positive = "1", levels = c(0,1)))[1],
+    .groups = "drop"
+  )
+
+# Merge AUCs back into pred_true
+pred_true <- left_join(pred_true, ind_auc, by = "subject")
+pred_true$auc_val = round(pred_true$auc_val,2)
+pred_true$auc_val[is.na(pred_true$auc_val)] = "-"
+
+pred_true$pred  = ifelse(pred_true$pred  > 0.5, 1, 0)
+
+# Add prediction correctness column
+pred_true <- pred_true %>%
+  mutate(prediction_correct = ifelse(true == pred, "Correct", "Incorrect")) # CHECK AGAIN IF PREDICTION IS 0/1 !!!
+
+plot = ggplot(pred_true, aes(day, subject, color = factor(true), shape = prediction_correct)) + 
+  geom_point(size = 2, alpha = 0.8) +
+  geom_text(aes(label = ifelse(!duplicated(subject), paste("AUC: ", auc_val), "")),
+            x = -Inf, hjust = 0, size = 3, color = "gray40") +  # AUC on the left side
+  scale_color_manual(values = c("0" = "#FFB5A7", "1" = "#40916C")) +
+  scale_shape_manual(values = c("Correct" = 15, "Incorrect" = 0)) +
+  labs(color = NULL, shape = NULL) +  # Removes legend titles
+  theme_minimal() +
+  theme(
+    panel.grid = element_blank(),
+    legend.position = "bottom",
+    axis.text.y = element_blank(),
+    axis.text.x = element_blank()     # Removes x-axis tick labels
+  ) +
+  xlim(-6, 75) +
+  labs(y = paste0("Within-Person"), x = "") +
+  ggtitle(paste("AUC evaluated over the whole dataset:", round(table$'AUC Overall', 2), "(Between-Person)")) +
+  theme(text = element_text('Futura'),axis.title=element_text(size=14)) 
+plot
+ggsave("WithinBetween_Example.png", 
+       width = 6100, height = 3000, units = "px", dpi = 800,plot = plot)
+
+```
+
 ```{r}
@@
 
-result <- read.csv("simulation_results.csv") # see Run_Simulation.R to investigate how results were created
+result <- read.csv("simulation_results_0.csv") # see Run_Simulation.R to investigate how results were created
 
-p1 <- ggplot(result[result$A == 0,], aes(x=icc, y=auc, color = icc_pred)) +
+p1 <- ggplot(result[result$A == 0,], aes(x=icc, y=auc, color = icc_pred)) + # color: icc_pred
   geom_point(alpha=0.4, size = 0.8) +
   theme_minimal() +
-  scale_colour_gradientn(colours = c("#2A363B","#83AF9B","#C8C8A9","#F9CDAD","#FC9D9A","#FE4365")) +
+  scale_colour_gradientn(colours = c("#1B4332", "#40916C", "#74C69D", "#B5E48C", "#FDE68A", "#FFB5A7")) +
   ylab("Prediction Performance (Overall AUC)") +
@@
   ggtitle("Bias in Performance Estimates (AUC) When No Relationship Exists: Expected AUC = 0.5") +
-  geom_hline(yintercept = 0.5, color = "#83AF9B", size = 0.8) +
   annotate("text", x = 0.07, y = 0.82, label =  "Greater bias when between person differences explain a \nlarger % of total variation in both outcome and predictors.",  hjust = 0) +
   geom_curve(
-    aes(x = 0.4, y = 0.85, xend = 0.68, yend = 0.87), 
+    aes(x = 0.2, y = 0.86, xend = 0.35, yend = 0.93), 
     arrow = arrow(length = unit(0.08, "inch")), 
@@
   ) + 
-  annotate("text", x = 0.38, y = 0.54, label = "Low bias if between person differences explain a small \n% of total variation in outcome, predictors, or both.",  hjust = 0) +
+  annotate("text", x = 0.38, y = 0.58, label = "Low bias if between person differences explain a small \n% of total variation in outcome, predictors, or both.",  hjust = 0) +
     geom_curve(
-    aes(x = 0.37, y = 0.54, xend = 0.33, yend = 0.51), 
+    aes(x = 0.37, y = 0.55, xend = 0.33, yend = 0.51), 
     arrow = arrow(length = unit(0.08, "inch")), 
@@
     curvature = 0.1
-  ) +
+  ) 
   # Adding a diagonal line to illustrate the linear relationship
-  geom_abline(slope = 0.5, intercept = 0.52, color = "#FE4365",  size = 0.8)
-
 ggsave("Figures/Figure4_Example2.jpg",
@@
 ```{r}
-############# Figure 5 #################
+############# Figure 5 BINARY #################
 ########################################
-result <- read.csv("simulation_results.csv")
+library(ggplot2)
+library(tidyr)
+library(dplyr)
+library(patchwork)
 
-result <- result[result$A != 0 & result$A != 0.8,]
+final_results_df_2_b = read.csv("final_results_df_2_b.csv")
+final_results_df_3_b = read.csv("final_results_df_3_b.csv")
+final_results_df_4_b = read.csv("final_results_df_4_b.csv")
+final_results_df_5_b = read.csv("final_results_df_5_b.csv")
 
-p3 <- ggplot(result, aes(x = icc, y = auc_c_individual - auc_individual)) +
-  geom_point(aes(color = ifelse(auc_c_individual - auc_individual > 0, "Above Zero", "Below or Equal to Zero")), 
-             alpha = 0.3, size = 0.8, show.legend = FALSE) +
-  scale_color_manual(values = c("Above Zero" = "#FC9D9A", "Below or Equal to Zero" = "#83AF9B")) +
-  theme_minimal() +
-  ylab("Performance Difference \n(centered - uncentered)") +
-  xlab("% of Total Variation Explained by Between Person Differences in the Outcome (ICC Outcome)") +
-  ggtitle("Within-Person Performance") +
-  geom_hline(yintercept = 0) + 
-  annotate("text", x = 0, y = 0.28, label = "Centered predictors perform better in most cases (within-person)", hjust = 0) +
-  annotate("text", x = 0, y = -0.08, label = "Uncentered predictors perform worse in most cases (within-person)", hjust = 0) +
-  ylim(c(-0.45,0.35)) +
-  geom_curve(
-    aes(x = -0.01, y = 0.01, xend = -0.01, yend = 0.25), 
-    arrow = arrow(length = unit(0.08, "inch")), 
-    size = 0.5, 
-    color = "#FC9D9A", 
-    curvature = 0
-  ) +
-    geom_curve(
-    aes(x = -0.01, y = 0.01, xend = -0.01, yend = -0.1), 
-    arrow = arrow(length = unit(0.08, "inch")), 
-    size = 0.5, 
-    color = "#83AF9B", 
-    curvature = 0
+plot_data = function(data_viz, spanpar){
+  # Reshape to long format
+  df_long <- data_viz %>%
+    pivot_longer(
+      cols = c(`shuffle_AUC.`, `movingwindowshuffle_AUC.`),
+      names_to = "AUC_type",
+      values_to = "AUC"
     )
 
-p4 <- ggplot(result, aes(x = icc, y = auc_c - auc)) +
-  geom_point(aes(color = ifelse(auc_c - auc > 0, "Above Zero", "Below or Equal to Zero")), 
-             alpha = 0.3, size = 0.8,show.legend = FALSE) +
-  scale_color_manual(values = c("Above Zero" = "#FC9D9A", "Below or Equal to Zero" = "#83AF9B")) +
+  p = ggplot(df_long, aes(x = icc_data, y = AUC, color = AUC_type)) +
+    geom_point(alpha = 0.8, size = 1) +
+    geom_smooth(se = FALSE, method = "loess", span = spanpar) +
     theme_minimal() +
-  ylab("Performance Difference \n(centered - uncentered)") +
-  xlab("% of Total Variation Explained by Between Person Differences in the Outcome (ICC Outcome)") +
-  ggtitle("") +
-  geom_hline(yintercept = 0) + 
-  annotate("text", x = 0, y = 0.28, label = "Centered predictors perform better when between person differences explain a small % of total variance (overall)", hjust = 0) +
-  annotate("text", x = 0.01, y = -0.39, label = "Uncentered predictors perform better when between person differences explain a large % of total variance (overall)", hjust = 0) +
-  ylim(c(-0.45,0.35)) +
-   ggtitle("Overall Performance (Within and Between-Person)")  +
-    geom_curve(
-    aes(x = -0.01, y = 0.01, xend = -0.01, yend = 0.27), 
-    arrow = arrow(length = unit(0.08, "inch")), 
-    size = 0.5, 
-    color = "#FC9D9A", 
-    curvature = 0
+    scale_color_manual(
+      values = c(
+        "shuffle_AUC." = "#40916C",
+        "movingwindowshuffle_AUC." = "#4D8FAC"
+      ), 
+      labels = c(
+        "shuffle_AUC." = "No true Relationship (Record-Wise)",
+        "movingwindowshuffle_AUC." = "No true Relationship (Moving-Window)"
+      )
     ) +
-    geom_curve(
-    aes(x = 0.8, y = 0.01, xend = 0.8, yend = -0.41), 
-    arrow = arrow(length = unit(0.08, "inch")), 
-    size = 0.5, 
-    color = "#83AF9B", 
-    curvature = 0
+    geom_hline(yintercept = 0.5, color = "grey", size = 0.8) +
+    labs(color = "") +
+    ylim(0.4, 0.95) +
+    ggtitle(paste0("N: ", max(data_viz$num_subjects),
+                   ", Timepoints: ", max(data_viz$num_samples),
+                   ",\nICC Predictors (mean): ", round(data_viz$icc_pred, 2))) +
+    ylab("Overall AUC") +
+    xlab("ICC Outcome")
+
+  return(p)
+}
+
+spanpar = 0.6
+
+p2 = plot_data(final_results_df_2_b, spanpar)
+p3 = plot_data(final_results_df_3_b, spanpar)
+p4 = plot_data(final_results_df_4_b, spanpar)
+p5 = plot_data(final_results_df_5_b, spanpar)
+
+# Combine all plots and collect legends at the bottom
+final_plot <- (p2 + p3 + p4 + p5) + 
+  plot_layout(ncol = 2, guides = "collect") & 
+  theme(legend.position = "bottom")
+
+# Add title and optional styling
+final_plot <- final_plot +
+  plot_annotation(
+    title = "Predictive Performance Shuffled Data Binary Outcomes",
+    theme = theme(
+      plot.title = element_text(size = 16, face = "bold", hjust = 0.5)
+    )
   )
 
-combined_plot <- (p3 / p4) +
-  plot_annotation(
-    title = "Performance difference: centered vs. not-centered",
-  ) 
-combined_plot
-
-ggsave(  "Figures/Figure5_Example4.jpg",
-  plot = combined_plot,
-  width = 10.2,
-  height = 5.5)
-
+print(final_plot)
+ggsave("Figures/Figure5_NullModelBinary.jpg",
+  plot = final_plot,
+  width = 7,
+  height = 6)
 ```
@@
 ```{r}
-############# Figure 6 #################
+############# Figure 6 Continiuous #################
 ########################################
-result <- read.csv("simulation_results.csv")
-result <- result[result$A != 0 & result$A != 0.8,]
-p1 <- ggplot(result, aes(x=auc_c, y=auc_c_individual, color = icc_pred)) +
-  geom_point(alpha=0.5, size = 1) +
-  #scale_size(range = c(.001, 6), name="") +
+library(ggplot2)
+library(tidyr)
+library(dplyr)
+library(patchwork)
+
+final_results_df_2 = read.csv("final_results_df_2.csv")
+final_results_df_3 = read.csv("final_results_df_3.csv")
+final_results_df_4 = read.csv("final_results_df_4.csv")
+final_results_df_5 = read.csv("final_results_df_5.csv")
+
+plot_data = function(data_viz, spanpar){
+  # Reshape to long format
+  df_long <- data_viz %>%
+    pivot_longer(
+      cols = c(shuffle_cor,movingwindowshuffle_cor),
+       names_to = "type",
+      values_to = "Correlation"
+    )
+
+  p = ggplot(df_long, aes(x = icc_data, y = Correlation, color = type)) +
+    geom_point(alpha = 0.8, size = 1) +
+    geom_smooth(se = FALSE, method = "loess", span = spanpar) +
     theme_minimal() +
-  scale_colour_gradientn(colours = c("#2A363B","#83AF9B","#C8C8A9","#F9CDAD","#FC9D9A","#FE4365")) +
-  ylab("AUC within-person centered") +
-  xlab("AUC centered") +
-  guides(col = guide_colourbar(title = "% of Total Variation Explained by \nBetween Person Differences \nin Predictors (ICC Predictor)")) +
-  geom_hline(yintercept = 0.5)   + 
-annotate("text", x = 0.5, y = 0.9, label = "Linear relationship between overall AUC \nand within-person AUC",  hjust = 0) +
-  geom_curve(
-    aes(x = 0.63, y = 0.87, xend = 0.75, yend = 0.8), 
-    arrow = arrow(length = unit(0.08, "inch")), 
-    size = 0.3, 
-    color = "gray20", 
-    curvature = -0.15
-  ) +
-  ggtitle("Uncentered Predictors")
+    scale_color_manual(values = c(
+      "shuffle_cor" = "#40916C",
+      "movingwindowshuffle_cor" = "#4D8FAC"
+    ), labels = c("shuffle_cor" = "No true Relationship (Record-Wise)", 
+                  "movingwindowshuffle_cor" ="No true Relationship (Moving-Window)"
+    )) +
+    labs(color = "") +
+    ggtitle(paste0("N: ", max(data_viz$num_subjects),
+                   ", Timepoints: ", max(data_viz$num_samples),
+                   ",\nICC Predictors (mean): ", round(data_viz$icc_pred, 2))) +
+    ylab("Overall Correlation") +
+    xlab("ICC Outcome") +
+    ylim(0, 0.7)
 
-p2 <- ggplot(result, aes(x=auc, y=auc_individual, color = icc_pred)) +
-  geom_point(alpha=0.5, size = 1) +
-  #scale_size(range = c(.001, 6), name="") +
-  theme_minimal() +
-  scale_colour_gradientn(colours = c("#2A363B","#83AF9B","#C8C8A9","#F9CDAD","#FC9D9A","#FE4365")) +
-  ylab("AUC within-person") +
-  xlab("AUC") +
-  guides(col = guide_colourbar(title = "% of Total Variation Explained by \nBetween Person Differences \nin Predictors (ICC Predictor)")) +
-  geom_hline(yintercept = 0.5) + 
-annotate("text", x = 0.5, y = 0.9, label = "If between person differences explain a large % of variation in predictors and outcome,\nthe overall AUC becomes high while the within-person AUC remains low.",  hjust = 0) +
-  geom_curve(
-    aes(x = 0.75, y = 0.82, xend = 0.9, yend = 0.6), 
-    arrow = arrow(length = unit(0.08, "inch")), 
-    size = 0.3, 
-    color = "gray20", 
-    curvature = 0.35
-  ) +
-  ggtitle("Centered Predictors")
+  return(p)
+}
 
-combined_plot <- (p2 / p1) +
+spanpar = 0.4
+
+p2 = plot_data(final_results_df_2, spanpar)
+p3 = plot_data(final_results_df_3, spanpar)
+p4 = plot_data(final_results_df_4, spanpar)
+p5 = plot_data(final_results_df_5, spanpar)
+
+# Combine the plots and collect the legend
+final_plot2 <- (p2 + p3 + p4 + p5) +
+  plot_layout(ncol = 2, guides = "collect") & 
+  theme(legend.position = "bottom")
+
+# Add title and optional styling
+final_plot2 <- final_plot2 +
   plot_annotation(
-    title = "Relationship Overall AUC and within-person AUC (A = 0.05 and A = 0.2)",
+    title = "Predictive Performance Shuffled Data Continuous Outcomes",
+    theme = theme(
+      plot.title = element_text(size = 16, face = "bold", hjust = 0.5)
     )
-combined_plot
+  )
 
-ggsave( "Figures/Figure6_Example3.jpg",
-  plot = combined_plot,
-  width = 11.5,
-  height = 6.5)
+print(final_plot2)
+ggsave("Figures/Figure6_NullModelContinious.jpg",
+  plot = final_plot2,
+  width = 7,
+  height = 6)
+
 ```
@@
 ```{r }
-############ Table 7 #############
+############ Figure 7 #############
 ##################################
 
+# Generating Outcome
+overall_prob_outcome <- 0.1  # Overall probability of 1 (for the entire dataset) (e.g., tracking depression 0.146)
+sd_outcome <- 0.2 # Controls variability BETWEEN different subject (nneds to be smaller than sqrt(overall_prob_outcome*(1-overall_prob_outcome)))
+
+# Generating Features
+A <- 0.05 # Relationship between features and outcome
+feature_std <- 0.1 # population level feature generating process
+B <- 0.8  # Cross-subject variability ("random effect") (added per participants for all timepoints)
+C <- 0.45 # Within-subject variability (added within participant for each timepoint)
+
+source("Simulation_Functions_notshiny.R")
+
+set.seed(12361488)
+features_sample <- create_data(n_features,90,150,A,feature_std,B,C,overall_prob_outcome,sd_outcome,0,0)
+
 # Merge the p-values with the original dataset for plotting
-data <- features_sample
+data <- features_sample[[2]]
 
@@
     y = "Outcome", 
-    title = paste("Within-person variability over time: \n88 participants (58.6%) do not experience within-person variability. ")
+    title = paste("Within-person variability over time: \n78 participants (52%) do not experience within-person variability. ")
   ) +
@@
 
-ggsave( "Figures/Figure7_Example_var.jpg",
-  plot = p2,
-  width = 8,
-  height = 7)
+# ggsave( "Figures/Figure7_Example_var.jpg",
+#   plot = p2,
+#   width = 8,
+#   height = 7)
 
@@
 ```{r }
-############ Table 5 #############
+############ Table 7 #############
 ##################################
@@
 
-test <- sim_1$ind_base[[1]] %>%
+test <- sim_4$ind_base[[1]] %>%
   group_by(subject) %>%
@@
   summarise(
-    auc_val = auc(roc(as.numeric(as.character(true)), as.numeric(as.character(pred)), quiet = TRUE))[1],
+    auc_val = pROC::auc(pROC::roc(as.numeric(as.character(true)), as.numeric(as.character(pred)), quiet = TRUE))[1],
     .groups = "drop" # Ungroup after summarizing
@@
 mean(test$auc_val)
+
+```
+
+```{r}
+############# Figure 9 BINARY #################
+########################################
+library(ggplot2)
+library(tidyr)
+library(dplyr)
+library(patchwork)
+
+final_results_df_2_b = read.csv("final_results_df_2_b.csv")
+final_results_df_3_b = read.csv("final_results_df_3_b.csv")
+final_results_df_4_b = read.csv("final_results_df_4_b.csv")
+final_results_df_5_b = read.csv("final_results_df_5_b.csv")
+
+plot_data = function(data_viz,spanpar){
+  # Reshape to long format
+  df_long <- data_viz %>%
+    pivot_longer(
+      cols = c(`real_AUC.random.intercept.only.`, `real_AUC.`,`movingwindow_AUC.`,`subjectwise_AUC.`,`movingwindow_AUC.random.intercept.only.`),
+      names_to = "AUC_type",
+      values_to = "AUC"
+    )
+  p = ggplot(df_long, aes(x = icc_data, y = AUC, color = AUC_type)) +
+    geom_point(alpha = 0.8, size = 1) +
+    geom_smooth(se = FALSE, method = "loess",span = spanpar) +
+    theme_minimal() +
+    scale_color_manual(values = c(
+      "real_AUC.random.intercept.only." ="#B5E48C" ,
+      "real_AUC." = "#74C69D",
+      "movingwindow_AUC." ="#9EC5E7",
+      "movingwindow_AUC.random.intercept.only." = "#A2D2FF",
+      "subjectwise_AUC." ="#FFB5A7"
+    ), 
+    labels = c(
+               "real_AUC.random.intercept.only."="Random Intercept only (Record-Wise)",
+               "real_AUC." ="Random Forest (Record-Wise) ",
+               "movingwindow_AUC." ="Random Forest (Moving-Window)",
+               "subjectwise_AUC." ="Random Forest (Subject-Wise)",
+               "movingwindow_AUC.random.intercept.only." = "Random Intercept only (Moving-Window)")) +
+    theme(legend.position = "none") +
+    geom_hline(yintercept = 0.5, color = "grey", size = 0.8) +
+    labs(color = "Model") +
+    ylim(0.4,0.95) +
+    ggtitle(paste0("N: ", max(data_viz$num_subjects),", ", "Timepoints: ", max(data_viz$num_samples),", ","\n",  "ICC Predictors (mean): ", round(data_viz$icc_pred,2))) +
+    ylab("Overall AUC") +
+    xlab("ICC Outcome")
+  return(p)
+}
+
+spanpar = 0.6
+
+p2 = plot_data(final_results_df_2_b, spanpar)
+p3 = plot_data(final_results_df_3_b, spanpar)
+p4 = plot_data(final_results_df_4_b, spanpar)
+p5 = plot_data(final_results_df_5_b, spanpar)
+
+# Combine all plots and collect legends at the bottom
+final_plot <- (p2 + p3 + p4 + p5) + 
+  plot_layout(ncol = 2, guides = "collect") & 
+  theme(legend.position = "right")
+
+# Add title and optional styling
+final_plot <- final_plot +
+  plot_annotation(
+    title = "Predictive Performance Baseline Comparison Binary Outcomes",
+    theme = theme(
+      plot.title = element_text(size = 16, face = "bold", hjust = 0.5)
+    )
+  )
+
+print(final_plot)
+ggsave("Figures/Figure9_BaselineBinary.jpg",
+  plot = final_plot,
+  width = 9,
+  height = 8)
+```
+
+```{r}
+############# Figure 10 Continiuous #################
+########################################
+library(ggplot2)
+library(tidyr)
+library(dplyr)
+library(patchwork)
+
+final_results_df_2 = read.csv("final_results_df_2.csv")
+final_results_df_3 = read.csv("final_results_df_3.csv")
+final_results_df_4 = read.csv("final_results_df_4.csv")
+final_results_df_5 = read.csv("final_results_df_5.csv")
+
+plot_data = function(data_viz,spanpar){
+  # Reshape to long format
+  df_long <- data_viz %>%
+    pivot_longer(
+    #  cols = c(`shuffle_mae`, `real_mae_base`, `real_mae`,`centered_mae`,`subjectwise_mae`),
+      cols = c(`real_cor_base`, `real_cor`,`movingwindow_cor`,`subjectwise_cor`,movingwindow_cor_base),
+      names_to = "type",
+      values_to = "Correlation"
+    )
+  p = ggplot(df_long, aes(x = icc_data, y = Correlation, color = type)) +
+    geom_point(alpha = 0.8, size = 1) +
+    geom_smooth(se = FALSE, method = "loess",span = spanpar) +
+    theme_minimal() +
+    scale_color_manual(values = c(
+      "real_cor_base" ="#B5E48C" ,
+      "real_cor" = "#74C69D",
+      "movingwindow_cor" = "#9EC5E7",
+      "movingwindow_cor_base" = "#A2D2FF",
+      "subjectwise_cor" ="#FFB5A7"
+    ), labels = c("real_cor_base"="Random Intercept only (Record-Wise)",
+                  "real_cor" ="Random Forest (Record-Wise)",
+                  "movingwindow_cor" ="Random Forest (Moving-Window) ",
+                  "subjectwise_cor" = "Random Forest (Subject-Wise)",
+                  "movingwindow_cor_base" = "Random Intercept only (Moving-Window)"
+    )) +
+    theme(legend.position = "none") +
+    labs(color = "Model") +
+   # xlim(0.1,1) +
+    ggtitle(paste0("N: ", max(data_viz$num_subjects),", ", "Timepoints: ", max(data_viz$num_samples),", ","\n",  "ICC Predictors (mean): ", round(data_viz$icc_pred,2))) +
+    ylab("Overall Correlation") +
+    xlab("ICC Outcome") +
+    ylim(0, 1)
+  return(p)
+}
+
+spanpar = 0.4
+
+p2 = plot_data(final_results_df_2, spanpar)
+p3 = plot_data(final_results_df_3, spanpar)
+p4 = plot_data(final_results_df_4, spanpar)
+p5 = plot_data(final_results_df_5, spanpar)
+
+# Combine all plots and collect legends at the bottom
+final_plot <- (p2 + p3 + p4 + p5) + 
+  plot_layout(ncol = 2, guides = "collect") & 
+  theme(legend.position = "right")
+
+# Add title and optional styling
+final_plot <- final_plot +
+  plot_annotation(
+    title = "Predictive Performance Baseline Comparison Continuous Outcomes",
+    theme = theme(
+      plot.title = element_text(size = 16, face = "bold", hjust = 0.5)
+    )
+  )
+
+print(final_plot)
+ggsave("Figures/Figure10_BaselineContinious.jpg",
+  plot = final_plot,
+  width = 9,
+  height = 8)
 
@@
 p7 <- ggplot(result, aes(x=icc, y=auc_value_base)) +
-  geom_point(alpha=0.5, size = 1.2, color = "#83AF9B") +
+  geom_point(alpha=0.5, size = 1.2, color = "#40916C") +
   #scale_size(range = c(.001, 6), name="") +
   theme_minimal() +
-  scale_colour_gradientn(colours = c("#83AF9B")) +
+  scale_colour_gradientn(colours = c("#40916C")) +
   ylab("Overall Performance (AUC)") +
@@
     size = 0.3, 
-    color = "gray20", 
+    color = "#1B4332", 
     curvature = 0.2

### 2025-06-10 EmpiricalExample.R

@@ -1,0 +1,1001 @@
+
+library(dplyr)
+library(lme4)
+library(tibble)
+library(dplyr)
+library(pROC)
+library(tidyr)
+library(ggplot2)
+library(purrr)
+library(randomForest)
+
+run_simulation_example_icc <- function(data = data, days_inlcuded = 50,participants_included = 30, icc_treshhold = 0, direction = ">"){
+  
+  source("Simulation_UploadData.R")
+  set.seed(1236)
+  
+  data = data  %>% group_by %>%
+    group_by(user_id) %>%
+    mutate(day = row_number())
+  
+  data <- data %>%
+    group_by(user_id) %>%
+    filter(n_distinct(day) >= days_inlcuded) %>%
+    ungroup()
+  data = data[data$day <= days_inlcuded,]
+  print(length(unique(data$user_id)))
+
+  sampled_ids <- sample(unique(data$user_id), size = participants_included, replace = FALSE)
+
+  data <- data %>%
+    filter(user_id %in% sampled_ids)
+
+  
+  results_list <- list()     # For summary stats & ICC
+  model_list   <- list()     # For model results on shuffled data
+  
+  
+  id_variable = "user_id"   
+  time_variable = "day"
+  n_features_upload = c("wantingAnything", "wantingIntensity", 
+                        "withOthers", "feelingBadToGood", "energy", 
+                        "thinkingOverAndOver", "lonely", "selfWorth", "appreciating", "stressed")
+  
+  binary_vars = c("wantedToFeel_pleasant", "wantedToFeel_control", "wantedToFeel_goodAboutMe", 
+                  "wantedToFeel_connected", "wantedToFeel_comfort", "wantedToFeel_calm", 
+                  "wantedToFeel_excited", "wantedToFeel_competent", "wantedToFeel_acknowledged", 
+                  "wantedToFeel_alert", "physically_hungry", 
+                  "physically_tired", "physically_uncomfortable", "physically_energized", 
+                  "physically_good", "feeling_angry", "feeling_anxious", 
+                  "feeling_awe", "feeling_compassion", "feeling_content", "feeling_enthusiastic", 
+                  "feeling_grateful", "feeling_guilty", "feeling_happy", "feeling_resentful", 
+                  "feeling_restless", "feeling_sad")
+  
+  binary_vars = c("wantedToFeel_pleasant")
+  
[…]

### 2025-06-10 EmpiricalExample_continious.R

@@ -1,0 +1,445 @@
+
+download.OSF.file <- function(GUID,Access_Token=NULL,file_name)
+{
+  require(httr)
+  require(rjson)
+  #search for file private/public status
+  GETurl <- paste0("https://api.osf.io/v2/files/",GUID)
+  req <- GET(GETurl, write_disk("test",overwrite=T))
+  json_data <- fromJSON(file = "test")
+  if (length(json_data$data) > 0){
+    req1 <- GET(json_data$data$links$download,
+                write_disk(file_name, overwrite = TRUE))
+    print(paste0("The file has been downloaded to your working directory as: ",
+                 file_name))
+  }
+  else if (length(Access_Token) == 1){
+    if (grepl("https://osf.io",Access_Token)==TRUE){
+      req1 <- GET(paste0("https://api.osf.io/v2/files/",GUID,"/",gsub(".*/","",Access_Token)),
+                  write_disk("test", overwrite = TRUE))
+      json_data <- fromJSON(file = "test")
+      if (length(json_data$data) > 0){
+        req1 <- GET(json_data$data$links$download,
+                    write_disk(file_name, overwrite = TRUE))
+        print(paste0("The file has been downloaded to your working directory as: ",
+                     file_name))
+      }
+      else{
+        print(json_data$errors[[1]]$detail[1])
+      }
+    }
+    else if (grepl("https://osf.io",Access_Token)==FALSE){
+      req1 <- GET(paste0("https://api.osf.io/v2/files/",GUID),
+                  write_disk("test", overwrite = TRUE),
+                  add_headers("Authorization" = paste0("Bearer ",Access_Token)))
+      json_data <- fromJSON(file = "test")
+      if (length(json_data$data) > 0){
+        req1 <- GET(json_data$data$links$download,
+                    write_disk(file_name, overwrite = TRUE),
+                    add_headers("Authorization" = paste0("Bearer ",Access_Token)))
+        print(paste0("The file has been downloaded to your working directory as: ",
+                     file_name))
+      }
+      else{
+        print(json_data$errors[[1]]$detail[1])
+      }
+    }
+    else{
+      print(json_data$errors[[1]]$detail[1])
+    }
+  }
+  else{
+    print(json_data$errors[[1]]$detail[1])
+  }
+}
[…]

### 2025-06-10 Run_Simulation.R

@@ -10,32 +10,87 @@ library(lme4)
 library(caret)
 library(pROC)
 
+source("Simulation_Functions_notshiny.R")
+
+# Basic Parameters from Study
+n_features <- 10  # Features
+n_samples <- 30  # Timepoints
+n_subjects <- 10 # Subjects
+
+# Generating Outcome
+overall_prob_outcome <- 0.2  # Overall probability of 1 (for the entire dataset) (e.g., tracking depression 0.146)
+sd_outcome <- 0.2 # Controls variability BETWEEN different subject (nneds to be smaller than sqrt(overall_prob_outcome*(1-overall_prob_outcome)))
+
+# Generating Features
+A <- 0.05 # Relationship between features and outcome
+feature_std <- 0.1 # population level feature generating process
+B <- 0.8  # Cross-subject variability ("random effect") (added per participants for all timepoints)
+C <- 0.2 # Within-subject variability (added within participant for each timepoint)
+
+time_effect = 0
+timevariability = 0
+
+
+features_sample <- create_data(n_features, n_samples, n_subjects, A, feature_std, B, C, overall_prob_outcome, sd_outcome, time_effect,timevariability)
+features_sample_Astd <- features_sample[[2]]
+features_sample <- features_sample[[1]]
+
+test = features_sample  %>%
+  group_by(subject) %>%
+  summarize(mean =mean(prob))
+
+mean(test$mean)
+sd(test$mean)
 ####################################################
 ################ Run Simulation ####################
 ####################################################
 source("Simulation_Functions_notshiny.R")
 
+# n_features <- 10 
+# n_samples <- 90  
+# n_subjects <- 150 
+# time_effect = FALSE
+# 
+# A <- 0 # Relationship between features and outcome, we also ran this for A = 0.05 and A = 0.2, and then merged the results
+# feature_std <- seq(0.1,0.5, by = 0.15) # population level feature generating process
+# 
+# # Initialize parameters
+# overall_prob_outcome <- seq(0.1, 0.9, by = 0.2)  
+# sd_outcome <- seq(0.05, 0.29, by = 0.1)            
+# B <- seq(0.1, 0.9, by = 0.1)       
+# C <-  seq(0.05,0.45, by = 0.1)   
+# 
+# time_effect <- c(0.1,0.2,0.4,0.6,0.8)
+# timevariability <- c(2,4,8,10)
+
+### Old
+
 n_features <- 10 
[…]

### 2025-06-10 Simulation_Functions_notshiny.R

@@ -2,13 +2,55 @@
 ### Simulation Functions #####
 ##############################
 
+### Helper Function ####
+
 estBetaParams <- function(mu, var) {
   alpha <- ((1 - mu) / var - 1 / mu) * mu ^ 2
   beta <- alpha * (1 / mu - 1)
   return(params = list(alpha = alpha, beta = beta))
 } #taken from https://stats.stackexchange.com/questions/12232/calculating-the-parameters-of-a-beta-distribution-using-the-mean-and-variance/12239#12239
 
-create_data <- function(n_features,n_samples,n_subjects,A,feature_std,B,C,overall_prob_outcome,sd_outcome,time_effect){
+
+#### Dependencies over time ####
+# The functions to create dependencies over time are adapted from 
+# Bringmann et al., "Changing dynamics: TV-AR models using generalized additive modeling"
+
+## Probability varies over time and follows a sinus function
+sine<-function(n_samples,time_effect,participant_mean, timevariability){ # TempDepen is the maximum absolute value of the function (time effect).  
+
+  genfun=rep(NA,(n_samples)) 
+  tt=1:(n_samples) #Defining a time parameter in order to create the sine function.
+  
+  genfun=(time_effect*participant_mean)*sin(timevariability*pi*tt/(n_samples)) + participant_mean #Here the actual sine function is created, we add TempDepen to have a value between 0 and 1       
+  
+  return(genfun)
+} 
+
+# ## AR 1 (still needs to be adapted)
+# 
+# AR1<-function(n_samples,beta_0,beta_1,sd_innovation, prob_ind){
+#   y<-rep(NA,n_samples) 
+# 
+#   y[1]<-rbinom(n_samples, size = 1, prob = prob_ind)
+# 
+#   for (t in 2:NT){ #In this for-loop the time points 2 to NT are created.
+#     #beta_0 is the intercept, beta_1 the autoregressive coefficient.
+#     #rnorm(1,mean=0,sd=sd_innovation) is the innovation or error drawn from a normal distribution with mean zero and standard deviation (sd)
+#     #sd_innovation (standard deviation of the innovation; 1 in this case, thus variance and sd are the same here).
+#     y[t]<-exp(beta_0*prob_ind+beta_1*y[t-1]+rnorm(1,mean=0,sd=sd_innovation)
[…]

### 2025-06-10 Simulation_UploadData.R

@@ -1,0 +1,758 @@
+##############################
+### Simulation Functions #####
+##############################
+
+shuffle_data <- function(data, subject_var, outcome_var) {
+  # Shuffle participant IDs:
+  unique_subjects <- unique(data[[subject_var]])
+  shuffled_subjects <- sample(unique_subjects)
+  mapping <- data.frame(old_subject = unique_subjects, new_subject = shuffled_subjects,
+                        stringsAsFactors = FALSE)
+
+  # Merge mapping with data and update the subject column:
+  data <- data %>%
+    left_join(mapping, by = setNames("old_subject", subject_var)) %>%
+    mutate(!!sym(subject_var) := new_subject) %>%
+    select(-new_subject)
+  
+  # Shuffle the outcome variable within each subject:
+  data <- data %>%
+    group_by(!!sym(subject_var)) %>%
+    mutate(!!sym(outcome_var) := sample(!!sym(outcome_var))) %>%
+    ungroup()
+  
+  return(data)
+}
+
+
+########################################################
+############## Cross Validation Strategies #############
+########################################################
+
+############# Subject Wise and Row Wise ################
+
+run_simulation_own <- function(features_sample,cv,n_bootstrap,testsize, seed = "12361488",n_features){
+  set.seed(seed)
+  acc <- numeric(n_bootstrap)
+  auc_value <- numeric(n_bootstrap)
+  acc_rep <- numeric(n_bootstrap)
+  true_list <- list()  # To store true_sw for each iteration
+  pred_list <- list() 
+  subject <- list() 
+  ind <- list()
+  
+  acc_base <- numeric(n_bootstrap)
+  auc_value_base <- numeric(n_bootstrap)
+  acc_rep_base <- numeric(n_bootstrap)
+  true_list_base <- list()  # To store true_sw for each iteration
+  pred_list_base <- list() 
+  ind_base <- list()
+  
+  for (i in 1:n_bootstrap) {
+    n_subjects  = length(unique(features_sample$subject))
+    n_samples  = length(unique(features_sample$time))
+
+    
+    #### Row wise and subject wise ####
+    #  Prepare training and testing sets
+    if(cv == "record-wise"){
+      samples_test <- sample(x =nrow(features_sample), size = floor(testsize * nrow(features_sample))) 
[…]

### 2025-06-10 Simulation_continous.R

@@ -1,0 +1,806 @@
+library("Metrics")
+library("TSrepr") #mape
+
+r2 <- function(true,pred){
+  rss <- sum((pred- true) ^ 2,na.rm = TRUE)  # residual sum of squares
+  tss <- sum((true- mean(true, na.rm = TRUE))^ 2,na.rm = TRUE)  # total sum of squares
+  rsq <- 1 - rss/tss
+  return(rsq)
+}
+
+##############################
+### Simulation Functions #####
+##############################
+
+shuffle_data <- function(data, subject_var, outcome_var) {
+  # Shuffle participant IDs:
+  unique_subjects <- unique(data[[subject_var]])
+  shuffled_subjects <- sample(unique_subjects) # shuffles order of subjects without replacement
+  mapping <- data.frame(old_subject = unique_subjects, new_subject = shuffled_subjects,
+                        stringsAsFactors = FALSE)
+  
+  # Merge mapping with data and update the subject column:
+  data <- data %>%
+    left_join(mapping, by = setNames("old_subject", subject_var)) %>%
+    mutate(!!sym(subject_var) := new_subject) %>%
+    select(-new_subject)
+  
+  # Shuffle the outcome variable within each subject:
+  data <- data %>%
+    group_by(!!sym(subject_var)) %>%
+    mutate(!!sym(outcome_var) := sample(!!sym(outcome_var))) %>%
+    ungroup()
+  
+  return(data)
+}
+
+
+########################################################
+############## Cross Validation Strategies #############
+########################################################
+
+############# Subject Wise and Row Wise ################
+
+run_simulation_own <- function(features_sample,cv,n_bootstrap,testsize, seed = "12361488",n_features){
+  set.seed(seed)
+  
+  cor <- numeric(n_bootstrap)
+  rsq <- numeric(n_bootstrap)
+  mape <- numeric(n_bootstrap)
+  mae <- numeric(n_bootstrap)
+  rae <- numeric(n_bootstrap)
+ 
+  true_list <- list()  # To store true_sw for each iteration
+  pred_list <- list() 
+  subject <- list() 
+  ind <- list()
+  
+  cor_base <- numeric(n_bootstrap)
+  rsq_base <- numeric(n_bootstrap)
+  mape_base <- numeric(n_bootstrap)
+  mae_base <- numeric(n_bootstrap)
+  rae_base <- numeric(n_bootstrap)
+  
[…]

### 2025-06-10 Paper_Results.Rmd
revisions

@@
 ## Pitfall 1: Within person performance Emprircal Example
-```{r }
-### Descriptives
-
-# Read the CSV files
-final_results_df_2_b <- read.csv("final_results_df_2_b.csv")
-final_results_df_3_b <- read.csv("final_results_df_3_b.csv")
-final_results_df_4_b <- read.csv("final_results_df_4_b.csv")
-final_results_df_5_b <- read.csv("final_results_df_5_b.csv")
-
-# Select specific columns
-cols_to_select1 <- c("outcome", "overall_prob_outcome", "sd_outcome",
-                  "icc_data")
-cols_to_select <- c("overall_prob_outcome", "sd_outcome",
-                   "icc_data")
-
-df2_selected <- final_results_df_2_b[, cols_to_select1]
-df3_selected <- final_results_df_3_b[, cols_to_select]
-df4_selected <- final_results_df_4_b[, cols_to_select]
-df5_selected <- final_results_df_5_b[, cols_to_select]
-
-# Rename columns to distinguish between dataframes (optional but recommended)
-names(df2_selected) <- paste0(names(df2_selected), "_df2")
-names(df3_selected) <- paste0(names(df3_selected), "_df3")
-names(df4_selected) <- paste0(names(df4_selected), "_df4")
-names(df5_selected) <- paste0(names(df5_selected), "_df5")
-
-# Combine the selected dataframes column-wise
-combined_df <- cbind(df2_selected, df3_selected, df4_selected, df5_selected)
-
-# View or write the combined data frame
-head(combined_df)
-
-combined_df <- combined_df %>%
-  mutate(across(where(is.numeric), ~ round(.x, 2)))
-write.csv(combined_df, "combined_des_binary.csv", row.names = FALSE)
-
-# Read the CSV files
-final_results_df_2 <- read.csv("final_results_df_2.csv")
-final_results_df_3 <- read.csv("final_results_df_3.csv")
-final_results_df_4 <- read.csv("final_results_df_4.csv")
-final_results_df_5 <- read.csv("final_results_df_5.csv")
-
-# Select specific columns
-cols_to_select1 <- c("outcome", "overall_prob_outcome", "sd_outcome",
-                  "icc_data")
-cols_to_select <- c("overall_prob_outcome", "sd_outcome",
-                   "icc_data")
-
-df2_selected <- final_results_df_2[, cols_to_select1]
-df3_selected <- final_results_df_3[, cols_to_select]
-df4_selected <- final_results_df_4[, cols_to_select]
-df5_selected <- final_results_df_5[, cols_to_select]
-
-# Rename columns to distinguish between dataframes (optional but recommended)
-names(df2_selected) <- paste0(names(df2_selected), "_df2")
-names(df3_selected) <- paste0(names(df3_selected), "_df3")
-names(df4_selected) <- paste0(names(df4_selected), "_df4")
-names(df5_selected) <- paste0(names(df5_selected), "_df5")
-
-# Combine the selected dataframes column-wise
-combined_df <- cbind(df2_selected, df3_selected, df4_selected, df5_selected)
-
-# View or write the combined data frame
-head(combined_df)
-
-combined_df <- combined_df %>%
-  mutate(across(where(is.numeric), ~ round(.x, 2)))
-write.csv(combined_df, "combined_des_continious.csv", row.names = FALSE)
-
-```
 
@@
   group_by(subject) %>%
-  mutate(y = lag(wantedToFeel_goodAboutMe, n = 1))
+  mutate(y = lag(wantedToFeel_goodAboutMe, n = 1)) # Lag outcome
 
@@
 rownames(table) <- c("row-wise","subject-wise","rolling-window")
-table
+round(table,2)
 ```
@@
 table
-```
-
-```{r }
-## Try out
-
-# Basic Parameters from Study
-n_features <- 10  # Features
-n_samples <- 200  # Timepoints
-n_subjects <- 15 # Subjects
-
-# Generating Outcome
-overall_prob_outcome <- 0.4  # Overall probability of 1 (for the entire dataset) (e.g., tracking depression 0.146)
-sd_outcome <- 0.3 # Controls variability BETWEEN different subject (nneds to be smaller than sqrt(overall_prob_outcome*(1-overall_prob_outcome)))
-
-# Generating Features
-A <- 0.02 # Relationship between features and outcome
-feature_std <- 0.1 # population level feature generating process
-B <- 0.8  # Cross-subject variability ("random effect") (added per participants for all timepoints)
-C <- 0.2 # Within-subject variability (added within participant for each timepoint)
-
-source("Simulation_Functions_notshiny.R")
-
-set.seed(12361488)
-features_sample <- create_data(n_features, n_samples, n_subjects, A, feature_std, B, C, overall_prob_outcome, sd_outcome, 0,0)
-
-features_sample <- features_sample[[1]]
-
-sim_1 <- run_simulation(features_sample,"row-wise",1, testsize = 0.3)
-
-table <- rbind(sim_1$overall_summary)
-table$'AUC Overall' <- c(sim_1$auc_value)
-
-table
-
-library(dplyr)
-library(ggplot2)
-
-pred_true = as.data.frame(sim_1$ind)
-
-pred_true <- pred_true %>%
-  arrange(subject, true) %>%
-  group_by(subject) %>%
-  mutate(day = row_number())
-
-library(pROC)
-library(dplyr)
-
-# Compute individual AUC
-ind_auc <- pred_true %>%
-  group_by(subject) %>%
-  filter(length(unique(true)) > 1) %>%
-  filter(sum(true == 1) > 0, sum(true == 0) > 0) %>%
-  summarise(
-    auc_val = pROC::auc(pROC::roc(response = true, predictor = pred, direction = "<",positive = "1", levels = c(0,1)))[1],
-    .groups = "drop"
-  )
-
-# Merge AUCs back into pred_true
-pred_true <- left_join(pred_true, ind_auc, by = "subject")
-pred_true$auc_val = round(pred_true$auc_val,2)
-pred_true$auc_val[is.na(pred_true$auc_val)] = "-"
-
-pred_true$pred  = ifelse(pred_true$pred  > 0.5, 1, 0)
-
-# Add prediction correctness column
-pred_true <- pred_true %>%
-  mutate(prediction_correct = ifelse(true == pred, "Correct", "Incorrect")) # CHECK AGAIN IF PREDICTION IS 0/1 !!!
-
-plot = ggplot(pred_true, aes(day, subject, color = factor(true), shape = prediction_correct)) + 
-  geom_point(size = 2, alpha = 0.8) +
-  geom_text(aes(label = ifelse(!duplicated(subject), paste("AUC: ", auc_val), "")),
-            x = -Inf, hjust = 0, size = 3, color = "gray40") +  # AUC on the left side
-  scale_color_manual(values = c("0" = "#FFB5A7", "1" = "#40916C")) +
-  scale_shape_manual(values = c("Correct" = 15, "Incorrect" = 0)) +
-  labs(color = NULL, shape = NULL) +  # Removes legend titles
-  theme_minimal() +
-  theme(
-    panel.grid = element_blank(),
-    legend.position = "bottom",
-    axis.text.y = element_blank(),
-    axis.text.x = element_blank()     # Removes x-axis tick labels
-  ) +
-  xlim(-6, 75) +
-  labs(y = paste0("Within-Person"), x = "") +
-  ggtitle(paste("AUC evaluated over the whole dataset:", round(table$'AUC Overall', 2), "(Between-Person)")) +
-  theme(text = element_text('Futura'),axis.title=element_text(size=14)) 
-plot
-ggsave("WithinBetween_Example.png", 
-       width = 6100, height = 3000, units = "px", dpi = 800,plot = plot)
-
 ```
@@
 ```
+
+Other things that were not presented in the paper (but used for a presentation):
+
+```{r }
+## Try out
+
+# Basic Parameters from Study
+n_features <- 10  # Features
+n_samples <- 200  # Timepoints
+n_subjects <- 15 # Subjects
+
+# Generating Outcome
+overall_prob_outcome <- 0.4  # Overall probability of 1 (for the entire dataset) (e.g., tracking depression 0.146)
+sd_outcome <- 0.3 # Controls variability BETWEEN different subject (nneds to be smaller than sqrt(overall_prob_outcome*(1-overall_prob_outcome)))
+
+# Generating Features
+A <- 0.02 # Relationship between features and outcome
+feature_std <- 0.1 # population level feature generating process
+B <- 0.8  # Cross-subject variability ("random effect") (added per participants for all timepoints)
+C <- 0.2 # Within-subject variability (added within participant for each timepoint)
+
+source("Simulation_Functions_notshiny.R")
+
+set.seed(12361488)
+features_sample <- create_data(n_features, n_samples, n_subjects, A, feature_std, B, C, overall_prob_outcome, sd_outcome, 0,0)
+
+features_sample <- features_sample[[1]]
+
+sim_1 <- run_simulation(features_sample,"row-wise",1, testsize = 0.3)
+
+table <- rbind(sim_1$overall_summary)
+table$'AUC Overall' <- c(sim_1$auc_value)
+
+table
+
+library(dplyr)
+library(ggplot2)
+
+pred_true = as.data.frame(sim_1$ind)
+
+pred_true <- pred_true %>%
+  arrange(subject, true) %>%
+  group_by(subject) %>%
+  mutate(day = row_number())
+
+library(pROC)
+library(dplyr)
+
+# Compute individual AUC
+ind_auc <- pred_true %>%
+  group_by(subject) %>%
+  filter(length(unique(true)) > 1) %>%
+  filter(sum(true == 1) > 0, sum(true == 0) > 0) %>%
+  summarise(
+    auc_val = pROC::auc(pROC::roc(response = true, predictor = pred, direction = "<",positive = "1", levels = c(0,1)))[1],
+    .groups = "drop"
+  )
+
+# Merge AUCs back into pred_true
+pred_true <- left_join(pred_true, ind_auc, by = "subject")
+pred_true$auc_val = round(pred_true$auc_val,2)
+pred_true$auc_val[is.na(pred_true$auc_val)] = "-"
+
+pred_true$pred  = ifelse(pred_true$pred  > 0.5, 1, 0)
+
+# Add prediction correctness column
+pred_true <- pred_true %>%
+  mutate(prediction_correct = ifelse(true == pred, "Correct", "Incorrect")) # CHECK AGAIN IF PREDICTION IS 0/1 !!!
+
+plot = ggplot(pred_true, aes(day, subject, color = factor(true), shape = prediction_correct)) + 
+  geom_point(size = 2, alpha = 0.8) +
+  geom_text(aes(label = ifelse(!duplicated(subject), paste("AUC: ", auc_val), "")),
+            x = -Inf, hjust = 0, size = 3, color = "gray40") +  # AUC on the left side
+  scale_color_manual(values = c("0" = "#FFB5A7", "1" = "#40916C")) +
+  scale_shape_manual(values = c("Correct" = 15, "Incorrect" = 0)) +
+  labs(color = NULL, shape = NULL) +  # Removes legend titles
+  theme_minimal() +
+  theme(
+    panel.grid = element_blank(),
+    legend.position = "bottom",
+    axis.text.y = element_blank(),
+    axis.text.x = element_blank()     # Removes x-axis tick labels
+  ) +
+  xlim(-6, 75) +
+  labs(y = paste0("Within-Person"), x = "") +
+  ggtitle(paste("AUC evaluated over the whole dataset:", round(table$'AUC Overall', 2), "(Between-Person)")) +
+  theme(text = element_text('Futura'),axis.title=element_text(size=14)) 
+plot
+# ggsave("WithinBetween_Example.png", 
+#        width = 6100, height = 3000, units = "px", dpi = 800,plot = plot)
+
+```

### 2025-06-10 EmpiricalExample.R

@@ -9,105 +9,10 @@ library(ggplot2)
 library(purrr)
 library(randomForest)
 
-run_simulation_example_icc <- function(data = data, days_inlcuded = 50,participants_included = 30, icc_treshhold = 0, direction = ">"){
-  
-  source("Simulation_UploadData.R")
-  set.seed(1236)
-  
-  data = data  %>% group_by %>%
-    group_by(user_id) %>%
-    mutate(day = row_number())
-  
-  data <- data %>%
-    group_by(user_id) %>%
-    filter(n_distinct(day) >= days_inlcuded) %>%
-    ungroup()
-  data = data[data$day <= days_inlcuded,]
-  print(length(unique(data$user_id)))
-
-  sampled_ids <- sample(unique(data$user_id), size = participants_included, replace = FALSE)
-
-  data <- data %>%
-    filter(user_id %in% sampled_ids)
-
-  
-  results_list <- list()     # For summary stats & ICC
-  model_list   <- list()     # For model results on shuffled data
-  
-  
-  id_variable = "user_id"   
-  time_variable = "day"
-  n_features_upload = c("wantingAnything", "wantingIntensity", 
-                        "withOthers", "feelingBadToGood", "energy", 
-                        "thinkingOverAndOver", "lonely", "selfWorth", "appreciating", "stressed")
-  
-  binary_vars = c("wantedToFeel_pleasant", "wantedToFeel_control", "wantedToFeel_goodAboutMe", 
-                  "wantedToFeel_connected", "wantedToFeel_comfort", "wantedToFeel_calm", 
-                  "wantedToFeel_excited", "wantedToFeel_competent", "wantedToFeel_acknowledged", 
-                  "wantedToFeel_alert", "physically_hungry", 
-                  "physically_tired", "physically_uncomfortable", "physically_energized", 
-                  "physically_good", "feeling_angry", "feeling_anxious", 
-                  "feeling_awe", "feeling_compassion", "feeling_content", "feeling_enthusiastic", 
-                  "feeling_grateful", "feeling_guilty", "feeling_happy", "feeling_resentful", 
-                  "feeling_restless", "feeling_sad")
-  
-  binary_vars = c("wantedToFeel_pleasant")
-  
-  non_binar <- names(data)[sapply(data, function(x) length(unique(na.omit(x))) > 2)]
-  
-  
[…]

### 2025-06-10 EmpiricalExample_continious.R

@@ -1,81 +1,16 @@
+library(dplyr)
+library(lme4)
+library(tibble)
+library(dplyr)
+library(pROC)
+library(tidyr)
+library(ggplot2)
+library(purrr)
+library(randomForest)
 
-download.OSF.file <- function(GUID,Access_Token=NULL,file_name)
-{
-  require(httr)
-  require(rjson)
-  #search for file private/public status
-  GETurl <- paste0("https://api.osf.io/v2/files/",GUID)
-  req <- GET(GETurl, write_disk("test",overwrite=T))
-  json_data <- fromJSON(file = "test")
-  if (length(json_data$data) > 0){
-    req1 <- GET(json_data$data$links$download,
-                write_disk(file_name, overwrite = TRUE))
-    print(paste0("The file has been downloaded to your working directory as: ",
-                 file_name))
-  }
-  else if (length(Access_Token) == 1){
-    if (grepl("https://osf.io",Access_Token)==TRUE){
-      req1 <- GET(paste0("https://api.osf.io/v2/files/",GUID,"/",gsub(".*/","",Access_Token)),
-                  write_disk("test", overwrite = TRUE))
-      json_data <- fromJSON(file = "test")
-      if (length(json_data$data) > 0){
-        req1 <- GET(json_data$data$links$download,
-                    write_disk(file_name, overwrite = TRUE))
-        print(paste0("The file has been downloaded to your working directory as: ",
-                     file_name))
-      }
-      else{
-        print(json_data$errors[[1]]$detail[1])
-      }
-    }
-    else if (grepl("https://osf.io",Access_Token)==FALSE){
-      req1 <- GET(paste0("https://api.osf.io/v2/files/",GUID),
-                  write_disk("test", overwrite = TRUE),
-                  add_headers("Authorization" = paste0("Bearer ",Access_Token)))
-      json_data <- fromJSON(file = "test")
-      if (length(json_data$data) > 0){
-        req1 <- GET(json_data$data$links$download,
-                    write_disk(file_name, overwrite = TRUE),
-                    add_headers("Authorization" = paste0("Bearer ",Access_Token)))
-        print(paste0("The file has been downloaded to your working directory as: ",
-                     file_name))
-      }
-      else{
-        print(json_data$errors[[1]]$detail[1])
-      }
[…]

### 2025-06-10 Simulation_Functions_notshiny.R

@@ -12,6 +12,7 @@ estBetaParams <- function(mu, var) {
 
 
 #### Dependencies over time ####
+# Not used in this paper but could be added as a future adaptation
 # The functions to create dependencies over time are adapted from 
 # Bringmann et al., "Changing dynamics: TV-AR models using generalized additive modeling"
 
@@ -25,25 +26,6 @@ sine<-function(n_samples,time_effect,participant_mean, timevariability){ # TempD
   
   return(genfun)
 } 
-
-# ## AR 1 (still needs to be adapted)
-# 
-# AR1<-function(n_samples,beta_0,beta_1,sd_innovation, prob_ind){
-#   y<-rep(NA,n_samples) 
-# 
-#   y[1]<-rbinom(n_samples, size = 1, prob = prob_ind)
-# 
-#   for (t in 2:NT){ #In this for-loop the time points 2 to NT are created.
-#     #beta_0 is the intercept, beta_1 the autoregressive coefficient.
-#     #rnorm(1,mean=0,sd=sd_innovation) is the innovation or error drawn from a normal distribution with mean zero and standard deviation (sd)
-#     #sd_innovation (standard deviation of the innovation; 1 in this case, thus variance and sd are the same here).
-#     y[t]<-exp(beta_0*prob_ind+beta_1*y[t-1]+rnorm(1,mean=0,sd=sd_innovation)
-#     
-#     
-#   }
-# 
-#   return(y)
-# }
 
 
 

### 2025-12-04 Paper_Results.Rmd
revision r1 AMPPS

@@
 library(patchwork)
+library(CorBin)
 ```
@@
 # Generating Features
-A <- 0.05 # Relationship between features and outcome
+A <- 0.05#1 # Relationship between features and outcome
 feature_std <- 0.1 # population level feature generating process
@@
 source("Simulation_Functions_notshiny.R")
+time_type = "None" #DCP, 1-dependent
+time_effect = 0
 
 set.seed(12361488)
-features_sample <- create_data(n_features,n_samples,n_subjects,A,feature_std,B,C,overall_prob_outcome,sd_outcome,0,0)
+features_sample <-create_data(n_features, n_samples, n_subjects, A, feature_std, B, C, overall_prob_outcome, sd_outcome, time_effect, time_type)
+#features_sample <- create_data(n_features,n_samples,n_subjects,A,feature_std,B,C,overall_prob_outcome,sd_outcome,0,0)
 features_sample <- features_sample[[1]]
@@
 table
+
+sub_id <- unique(features_sample$subject)
+
+auc_value_ind  <- numeric(length(sub_id))
+auc_value_base <- numeric(length(sub_id))
+sub_number     <- sub_id  # initialize
+
+for (k in seq_along(sub_id)) {
+
+  idx <- features_sample$subject == sub_id[k]
+  y_k <- features_sample$y[idx]
+
+  # Only proceed if conditions are met
+  if (sum(y_k) > 0 && length(unique(y_k)) > 1) {
+    message("Processing subject ", k, " (ID: ", sub_id[k], ")")
+
+    tryCatch(
+      {
+        res <- run_simulation_slidingwindow(
+          features_sample[idx, ],
+          1,
+          windowsize = 14
+        )
+
+        auc_value_ind[k]  <- res$auc_value
+        auc_value_base[k] <- res$auc_value_base
+      },
+      error = function(e) {
+        warning("Simulation failed for subject ", sub_id[k], ": ", e$message)
+        auc_value_ind[k]  <- NA
+        auc_value_base[k] <- NA
+      }
+    )
+  } else {
+    auc_value_ind[k]  <- NA
+    auc_value_base[k] <- NA
+  }
+}
+
+mean(auc_value_ind, na.rm = TRUE)
+
+sum(!is.na(auc_value_ind))
+
+sum(auc_value_ind[!is.na(auc_value_ind)] != 0)
+
+# Table A.1.1
+
+time_type = "1-dependent" 
+time_effect = 0.25
+
+set.seed(12361488)
+rm(features_sample)
+features_sample <-create_data(n_features, n_samples, n_subjects, A, feature_std, B, C, overall_prob_outcome, sd_outcome, time_effect, time_type)
+features_sample <- features_sample[[1]]
+
+sim_1 <- run_simulation(features_sample,"row-wise",1, testsize = 0.3)
+sim_2 <- run_simulation_centering(features_sample,"row-wise",1,testsize = 0.3)
+sim_3 <- run_simulation(features_sample,"subject-wise",1, testsize = 0.3)
+sim_4 <- run_simulation_slidingwindow(features_sample,1,windowsize = 14)
+
+table <- rbind(sim_1$overall_summary,sim_3$overall_summary, sim_4$overall_summary)
+table$'AUC Overall' <- c(sim_1$auc_value,sim_3$auc_value,sim_4$auc_value)
+colnames(table) <-  c("Mean AUC within-person:", "SD AUC within-person:", "% of AUC > 0.5 within-person:", "N included within-person:", "AUC Overall")
+rownames(table) <- c("row-wise","subject-wise","rolling-window")
+table <- round(table[,c(5,1,2,3,4)], 2)
+table
+
+#Table A.1.2
+time_type = "DCP"
+time_effect =  seq(0.9, 0.01, length.out = (n_samples-1)) #0.2
+
+set.seed(12361488)
+rm(features_sample)
+features_sample <-create_data(n_features, n_samples, n_subjects, A, feature_std, B, C, overall_prob_outcome, sd_outcome, time_effect, time_type)
+features_sample <- features_sample[[1]]
+
+sim_1 <- run_simulation(features_sample,"row-wise",1, testsize = 0.3)
+sim_2 <- run_simulation_centering(features_sample,"row-wise",1,testsize = 0.3)
+sim_3 <- run_simulation(features_sample,"subject-wise",1, testsize = 0.3)
+sim_4 <- run_simulation_slidingwindow(features_sample,1,windowsize = 14)
+
+table <- rbind(sim_1$overall_summary,sim_3$overall_summary, sim_4$overall_summary)
+table$'AUC Overall' <- c(sim_1$auc_value,sim_3$auc_value,sim_4$auc_value)
+colnames(table) <-  c("Mean AUC within-person:", "SD AUC within-person:", "% of AUC > 0.5 within-person:", "N included within-person:", "AUC Overall")
+rownames(table) <- c("row-wise","subject-wise","rolling-window")
+table <- round(table[,c(5,1,2,3,4)], 2)
+table
+
+# Table A.1.3
+set.seed(12361488)
+features_sample <-create_data(n_features, n_samples, n_subjects, A, feature_std = 2, B, C, overall_prob_outcome, sd_outcome, 0, "None")
+features_sample <- features_sample[[1]]
+
+sim_1 <- run_simulation(features_sample,"row-wise",1, testsize = 0.3)
+sim_2 <- run_simulation_centering(features_sample,"row-wise",1,testsize = 0.3)
+sim_3 <- run_simulation(features_sample,"subject-wise",1, testsize = 0.3)
+sim_4 <- run_simulation_slidingwindow(features_sample,1,windowsize = 14)
+
+table <- rbind(sim_1$overall_summary,sim_3$overall_summary, sim_4$overall_summary)
+table$'AUC Overall' <- c(sim_1$auc_value,sim_3$auc_value,sim_4$auc_value)
+colnames(table) <-  c("Mean AUC within-person:", "SD AUC within-person:", "% of AUC > 0.5 within-person:", "N included within-person:", "AUC Overall")
+rownames(table) <- c("row-wise","subject-wise","rolling-window")
+table <- round(table[,c(5,1,2,3,4)], 2)
+table
+
 ```
@@
 result <- read.csv("simulation_results_0.csv") # see Run_Simulation.R to investigate how results were created
+title = "Bias in Performance Estimates (AUC) When No Relationship Exists: Expected AUC = 0.5"
 
-p1 <- ggplot(result[result$A == 0,], aes(x=icc, y=auc, color = icc_pred)) + # color: icc_pred
+p1 <- ggplot(result, aes(x=icc, y=auc, color = icc_pred)) + # color: icc_pred
   geom_point(alpha=0.4, size = 0.8) +
@@
   guides(col = guide_colourbar(title = "% of Total Variation Explained by \nBetween Person Differences \nin Predictors (ICC Predictor)")) +
-  ggtitle("Bias in Performance Estimates (AUC) When No Relationship Exists: Expected AUC = 0.5") +
+  ggtitle(title) +
   annotate("text", x = 0.07, y = 0.82, label =  "Greater bias when between person differences explain a \nlarger % of total variation in both outcome and predictors.",  hjust = 0) +
@@
   ) 
-  # Adding a diagonal line to illustrate the linear relationship
-ggsave("Figures/Figure4_Example2.jpg",
-  plot = p1,
-  width = 10.2,
-  height = 4.5)
+#   # Adding a diagonal line to illustrate the linear relationship
+# ggsave("Figures/Figure4_Example2.jpg",
+#   plot = p1,
+#   width = 10.2,
+#   height = 4.5)
 
@@
 sd(result$auc_individual[result$A == 0])
+
+result <- read.csv("simulation_results_time_A11.csv") # autoregressive effect in predictors
+title = "Robustness Check 1: Bias in Performance Estimates (AUC) When No Relationship Exists: Expected AUC = 0.5"
+
+result <- read.csv("simulation_results_time_A12.csv") # autoregressive effect in predictors
+title = "Robustness Check 2: Bias in Performance Estimates (AUC) When No Relationship Exists: Expected AUC = 0.5"
+
+result <- read.csv("simulation_results_time_A13.csv") # autoregressive effect in predictors
+result <- result[result$feature_std ==2,]
+
+title = "Robustness Check 3: Bias in Performance Estimates (AUC) When No Relationship Exists: Expected AUC = 0.5"
+
+p1 <- ggplot(result, aes(x=icc, y=auc, color = icc_pred)) + # color: icc_pred
+  geom_point(alpha=0.4, size = 0.8) +
+  theme_minimal() +
+  scale_colour_gradientn(colours = c("#1B4332", "#40916C", "#74C69D", "#B5E48C", "#FDE68A", "#FFB5A7")) +
+  ylab("Prediction Performance (Overall AUC)") +
+  xlab("% of Total Variation Explained by Between Person Differences in the Outcome \n(ICC Outcome)") +
+  guides(col = guide_colourbar(title = "% of Total Variation Explained by \nBetween Person Differences \nin Predictors (ICC Predictor)")) +
+  ggtitle(title)
+
+  # Adding a diagonal line to illustrate the linear relationship
+ggsave("Figures/Figure_A13.jpg",
+  plot = p1,
+  width = 10.2,
+  height = 4.5)
+
+p1
 
@@
       values = c(
-        "shuffle_AUC." = "#40916C",
-        "movingwindowshuffle_AUC." = "#4D8FAC"
+        "shuffle_AUC." = "#c7835a",
+        "movingwindowshuffle_AUC." = "#58538f"
       ), 
@@
     scale_color_manual(values = c(
-      "shuffle_cor" = "#40916C",
-      "movingwindowshuffle_cor" = "#4D8FAC"
+      "shuffle_cor" = "#c7835a",
+      "movingwindowshuffle_cor" = "#58538f"
     ), labels = c("shuffle_cor" = "No true Relationship (Record-Wise)", 
@@
     xlab("ICC Outcome") +
-    ylim(0, 0.7)
+    ylim(-0.1, 0.7)
 
@@
   ) +
-  geom_point(color = "#FC9D9A", size = 0.5) +                       
+  geom_point(color = "#c7835a", size = 0.5) +                       
   facet_wrap(~subject) +               
@@
 ##################################
-sim_1$auc_value_base # row-wise
-sim_4$auc_value_base # rolling window
+round(sim_1$auc_value_base,2) # row-wise
+round(sim_4$auc_value_base,2) # rolling window
+sim_3$auc_value_base # subject-wise
 
@@
 test <- sim_4$ind_base[[1]] %>%
+  group_by(subject) %>%
+  filter(length(unique(true)) > 1) %>%
+  filter(sum(true == 1) > 0, sum(true == 0) > 0) %>%
+  summarise(
+    auc_val = pROC::auc(pROC::roc(as.numeric(as.character(true)), as.numeric(as.character(pred)), quiet = TRUE))[1],
+    .groups = "drop" # Ungroup after summarizing
+  )
+
+mean(test$auc_val)
+
+test <- sim_3$ind_base[[1]] %>%
   group_by(subject) %>%
@@
     scale_color_manual(values = c(
-      "real_AUC.random.intercept.only." ="#B5E48C" ,
-      "real_AUC." = "#74C69D",
-      "movingwindow_AUC." ="#9EC5E7",
-      "movingwindow_AUC.random.intercept.only." = "#A2D2FF",
-      "subjectwise_AUC." ="#FFB5A7"
+      "real_AUC.random.intercept.only." ="#c7835a" ,
+      "real_AUC." = "#9c5435",
+      "movingwindow_AUC." ="#877fbf",
+      "movingwindow_AUC.random.intercept.only." = "#58538f",
+      "subjectwise_AUC." ="#40916c"
     ), 
@@
     scale_color_manual(values = c(
-      "real_cor_base" ="#B5E48C" ,
-      "real_cor" = "#74C69D",
-      "movingwindow_cor" = "#9EC5E7",
-      "movingwindow_cor_base" = "#A2D2FF",
-      "subjectwise_cor" ="#FFB5A7"
+      "real_cor_base" ="#c7835a" ,
+      "real_cor" = "#9c5435",
+      "movingwindow_cor" = "#877fbf",
+      "movingwindow_cor_base" = "#58538f",
+      "subjectwise_cor" ="#40916c"
     ), labels = c("real_cor_base"="Random Intercept only (Record-Wise)",
@@
 result = read.csv("simulation_results_baseline.csv")
+title = paste("Performance of a Random Intercept Only Model (No Predictors Included)")
+
+result = read.csv("simulation_results_baseline_A21.csv")
+title = paste("Robustness Check 1: Performance of a Random Intercept Only Model (No Predictors Included)")
+
+result = read.csv("simulation_results_baseline_A22.csv")
+title = paste("Robustness Check 2: Performance of a Random Intercept Only Model (No Predictors Included)")
+
 p7 <- ggplot(result, aes(x=icc, y=auc_value_base)) +
@@
   guides(col = guide_colourbar()) +
-  ggtitle(paste("Performance of a Random Intercept Only Model (No Predictors Included)")) +
+  ggtitle(title) +
   geom_hline(yintercept = 0.5) +
@@
 
+ggsave(  "Figures/FigureA22.jpg",
+  plot = p7,
+  width = 10,
+  height = 5)
+
 ```
@@
 Other things that were not presented in the paper (but used for a presentation):
+
+```{r }
+# --- PARAMETERS ---
+a_values <- c(0.00, 0.05)        # outcome effect
+b_values <- c(0.1, 0.5, 0.9)     # cross-subject variability
+c_values <- c(0.05, 0.25, 0.45)  # within-subject variability
+d <- 0.1                         # noise
+
+# --- COMPUTE SNR ---
+results <- expand.grid(a = a_values, b = b_values, c = c_values)
+
+results$var_signal <- results$a^2
+results$var_noise  <- results$b^2 + results$c^2 + d^2
+results$SNR_linear <- results$var_signal / results$var_noise
+results$SNR_dB     <- ifelse(results$SNR_linear > 0,
+                             10 * log10(results$SNR_linear),
+                             -Inf)
+
+# --- SHOW RESULTS ---
+print(round(results, 6))
+
+b_grid <- seq(0.1, 0.9, by = 0.05)
+c_grid <- seq(0.05, 0.45, by = 0.05)
+a <- 0.05
+
+grid <- expand.grid(b = b_grid, c = c_grid)
+grid$SNR_linear <- a^2 / (grid$b^2 + grid$c^2 + d^2)
+grid$SNR_dB <- 10 * log10(grid$SNR_linear)
+
+summary(grid$SNR_linear)
+summary(grid$SNR_dB)
+
+# Optional: visualize as heatmap
+library(ggplot2)
+ggplot(grid, aes(x = b, y = c, fill = SNR_linear)) +
+  geom_tile(color = "white") +
+  scale_fill_viridis_c(option = "plasma") +
+  labs(title = "SNR (dB) across b and c values",
+       x = "Cross-subject variability (b)",
+       y = "Within-subject variability (c)",
+       fill = "SNR (dB)") +
+  theme_minimal(base_size = 14)
+
+result <- read.csv("signal_to_noise.csv")
+mean(result$R2_marginal)
+median(result$R2_marginal)
+min(result$R2_marginal)
+max(result$R2_marginal)
+
+```
 

### 2025-12-04 Before Revisions/Simulation_Functions_notshiny.R

@@ -1,0 +1,555 @@
+##############################
+### Simulation Functions #####
+##############################
+
+estBetaParams <- function(mu, var) {
+  alpha <- ((1 - mu) / var - 1 / mu) * mu ^ 2
+  beta <- alpha * (1 / mu - 1)
+  return(params = list(alpha = alpha, beta = beta))
+} #taken from https://stats.stackexchange.com/questions/12232/calculating-the-parameters-of-a-beta-distribution-using-the-mean-and-variance/12239#12239
+
+create_data <- function(n_features,n_samples,n_subjects,A,feature_std,B,C,overall_prob_outcome,sd_outcome,time_effect){
+  
+  y <- vector("numeric", length = n_subjects * n_samples) 
+  subject_id <- rep(1:n_subjects, each = n_samples)  # Repeat subject IDs for each timepoint
+  time_id <- rep(1:n_samples, times = n_subjects)  # Repeat timepoints for each subject
+  data <- data.frame(subject = subject_id, time = time_id, y = y)
+  
+  # Simulate Outcome
+  para <- estBetaParams(overall_prob_outcome,sd_outcome^2)
+  prob_l <- rbeta(n=n_subjects ,shape1 = para$alpha, shape2 = para$beta) # sample probability for each subject
+  
+  print(paste("Mean prob:",mean(prob_l)))
+  print(paste("SD prob:",sd(prob_l)))
+  
+  prob_l = rep(prob_l, each = n_samples) 
+  
+  if (time_effect) {
+    #Can be added in the future
+  } else {
+    prob_matrix = prob_l
+  }
+  
+  data$y <- rbinom(n_subjects * n_samples, size = 1, prob = prob_matrix) #sample 0 or 1 for each subject & timepoint, prob_matrix indicates the probability for each subject
+  data$A = data$y
+
+  # Simulate Features (adapted from Saeb et al. and https://aosmith.rbind.io/2018/04/23/simulate-simulate-part-2/)
+  data$A[data$y == 0] = -A # relationship to outcome
+  data$A[data$y == 1] = A
+  
+  features_sample <- list()
+  
+  features_sample_ABCstd <- as.data.frame(array(0, dim = c(n_subjects * n_samples, n_features)))
+  features_sample_ABCstd <- cbind(features_sample_ABCstd,data)
+  
+  features_sample_Astd <- as.data.frame(array(0, dim = c(n_subjects * n_samples, n_features)))
+  features_sample_Astd <- cbind(features_sample_Astd,data)
+  
[…]

### 2025-12-04 Run_Simulation.R
revisions AMPPS r1

@@ -9,31 +9,102 @@ library(dplyr)
 library(lme4)
 library(caret)
 library(pROC)
+library(CorBin)
+library(tidyr)
 
+
+
+source("Simulation_continous.R")
 source("Simulation_Functions_notshiny.R")
+
 
 # Basic Parameters from Study
 n_features <- 10  # Features
-n_samples <- 30  # Timepoints
-n_subjects <- 10 # Subjects
+n_samples <- 90  # Timepoints
+n_subjects <- 150 # Subjects
 
 # Generating Outcome
-overall_prob_outcome <- 0.2  # Overall probability of 1 (for the entire dataset) (e.g., tracking depression 0.146)
-sd_outcome <- 0.2 # Controls variability BETWEEN different subject (nneds to be smaller than sqrt(overall_prob_outcome*(1-overall_prob_outcome)))
+overall_prob_outcome <- 0.5  # 0.5 | Overall probability of 1 (for the entire dataset) (e.g., tracking depression 0.146)
+sd_outcome <- 0.25 # 0| Controls variability BETWEEN different subject (needs to be smaller than sqrt(overall_prob_outcome*(1-overall_prob_outcome)))
 
 # Generating Features
-A <- 0.05 # Relationship between features and outcome
+A <- 0.1 # Relationship between features and outcome
 feature_std <- 0.1 # population level feature generating process
-B <- 0.8  # Cross-subject variability ("random effect") (added per participants for all timepoints)
-C <- 0.2 # Within-subject variability (added within participant for each timepoint)
+B <- 0.8 #0.9  # Cross-subject variability ("random effect") (added per participants for all timepoints)
+C <- 0.45 #0.45 # Within-subject variability (added within participant for each timepoint)
 
-time_effect = 0
-timevariability = 0
+time_type = "DCP" #DCP, 1-dependent
+time_effect = 0.2
+time_effect =  seq(0.9, 0.01, length.out = (n_samples-1)) #0.2
+
+time_effect = c(rep(0.5,n_samples-1))
+
+#time_effect = c(rep(0.9,45),rep(0,44))
 
 
-features_sample <- create_data(n_features, n_samples, n_subjects, A, feature_std, B, C, overall_prob_outcome, sd_outcome, time_effect,timevariability)
+time_effect = NULL
+# 
+# lags <- c(1:(n_samples-1))
[…]

### 2025-12-04 Simulation_Functions_notshiny.R

@@ -2,6 +2,7 @@
 ### Simulation Functions #####
 ##############################
 
+library(CorBin)
 ### Helper Function ####
 
 estBetaParams <- function(mu, var) {
@@ -15,24 +16,39 @@ estBetaParams <- function(mu, var) {
 # Not used in this paper but could be added as a future adaptation
 # The functions to create dependencies over time are adapted from 
 # Bringmann et al., "Changing dynamics: TV-AR models using generalized additive modeling"
+# 
+# ## Probability varies over time and follows a sinus function
+# sine<-function(n_samples,time_effect,participant_mean, timevariability){ # TempDepen is the maximum absolute value of the function (time effect).  
+# 
+#   genfun=rep(NA,(n_samples)) 
+#   tt=1:(n_samples) #Defining a time parameter in order to create the sine function.
+#   
+#   genfun=(time_effect*participant_mean)*sin(timevariability*pi*tt/(n_samples)) + participant_mean #Here the actual sine function is created, we add TempDepen to have a value between 0 and 1       
+#   
+#   return(genfun)
+# } 
 
-## Probability varies over time and follows a sinus function
-sine<-function(n_samples,time_effect,participant_mean, timevariability){ # TempDepen is the maximum absolute value of the function (time effect).  
 
-  genfun=rep(NA,(n_samples)) 
-  tt=1:(n_samples) #Defining a time parameter in order to create the sine function.
-  
-  genfun=(time_effect*participant_mean)*sin(timevariability*pi*tt/(n_samples)) + participant_mean #Here the actual sine function is created, we add TempDepen to have a value between 0 and 1       
-  
-  return(genfun)
-} 
+# exchange: all variables share same correlation, rho = amount of that correlation
+# DCP: local correlation decay, nearby variables have more correlation (e.g., rho = seq(0.3, 0.05, length.out = 9))
+# 1 - dependent: only adjacent variables are correlated (Dejonckere chapter, ar parameter varied a lot)
+# General: lag structure can be specified, rho is a list
+
[…]
